### 2012-07-01 paper.tex
BLD Start new paper for mahotas
Recent discussion on mailing=list has convinced me this is necessary

\documentclass{article}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{python}
\usepackage[%
    natbib=true,%
    backend=biber,%
    backref=true,%
    citestyle=authoryear-comp,%
    bibstyle=authoryear,%
    maxbibnames=24,%
    maxcitenames=2,%
]{biblatex}

\hypersetup{%
    bookmarks=true,
    colorlinks=true,
    pdfauthor={Luis Pedro Coelho},
    citecolor=pondgreen,
    urlcolor=toastedchilipowder,
    linkcolor=toastedchilipowder,
}
\let\code\texttt
\addbibresource{references.bib}

\title{Mahotas: Open source software for scriptable computer vision}
\author{Luis Pedro Coelho}

\begin{document}
\maketitle

\section*{Abstract}
\section{Introduction}
\printbibliography
\end{document}

### 2012-07-01 build.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/zsh
+
+function build() {
+  input=$1
+  mkdir -p .$input.tex_files
+  cd .$input.tex_files
+  TEXINPUTS=.:..:../images/:../figures/: xelatex $input
+  BSTINPUTS=:..: BIBINPUTS=:..:.: biber $input
+  cp $input.pdf ..
+  cd ..
+}
+build paper

### 2012-07-04 paper.tex
TXT Copy over thesis section
In the thesis, this was an appendix

@@
 }
+\newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
 \let\code\texttt
@@
 \section{Introduction}
+
+Mahotas is a computer vision library for Python. It operates on numpy arrays,
+but its inner loops are implemented in \cpp{} for speed and ease of
+implementation.
+
+\subsection{Functionality and Interface}
+
+The interface is a procedural interface, with most functions working
+independently of each other (there is code sharing at the implementation
+level).
+
+\begin{description}
+\item[features] Some feature descriptors. In particular, Haralick texture
+features, Zernike moments, local binary patterns, threshold adjacency
+statistics.
+\item[morphological functions] Erosion and dilation, as well as some more
+complex operations.
+\item[watershed] seeded watershed.
+\item[thresholding] some thresholding methods.
+\item[polygon operations] convex hull, polygon drawing.
+\item[\textsc{surf}] Speeded-up Robust Features.
+\end{description}
+
+There are a few interface conventions which apply to many functions.
+
+When meaningful, a structuring element is used to define neighbourhoods or
+adjacency relationships (some functions which use this concept are: watershed,
+erosion, and median filtering). Generally, the default is to use a $3 \times 3$
+cross as the default if no structuring filter is given (the exception to this
+rule is the median filter, where the default is a $3 \times 3$ square).
+
+Often, functions take an argument named \code{output} where the output will be
+stored. This argument is often much more restricted in type. In particular, it
+must often be a contiguous array.\footnote{Numpy supports non-contiguous
+arrays, which are most often slices into other, larger, contiguous arrays
+(e.g., given a $128 \times 128$ contiguous array, one can build a $64 \times
+128$ non-contiguous array by taking every other row).} Since this is a
+performance feature, it is natural that the interface is less flexible
+(accessing a contiguous array is much more efficient than a non-contiguous
+one).
+
+\subsection{Implementation}
+
+Mahotas is written in \cpp, but almost always, the user calls a Python function
+which checks types and then calls the internal function. This is slightly
+slower, but it is easier to develop this way.
+
+The main reason that mahotas is in \cpp{} (and not in pure C) is to use templates.
+Almost \cpp{} functionality is split across 2~functions:
+
+\begin{enumerate}
+\item A \code{py\_function} which uses the Python C~API to get arguments and
+check them. This is almost always pure~C.
+\item A template \code{function<dtype>} which works for the type \code{dtype}
+performing the actual operation.
+\end{enumerate}
+
+So, for example, this is how \emph{erode} is implemented. \code{py\_erode}
+consists mostly of boiler-plate code:
+
+\begin{cplusplus}
+PyObject* py_erode(PyObject* self, PyObject* args) {
+    PyArrayObject* array;
+    PyArrayObject* Bc;
+    if (!PyArg_ParseTuple(args,"OO", &array, &Bc)) {
+        return NULL;
+    }
+    PyArrayObject* res_a = (PyArrayObject*)PyArray_SimpleNew(
+                                array->nd,
+                                array->dimensions,
+                                PyArray_TYPE(array));
+    if (!res_a) return NULL;
+    PyArray_FILLWBYTE(res_a, 0);
+    switch(PyArray_TYPE(array)) {
+#define HANDLE(type) \
+    erode<type>(numpy::aligned_array<type>(res_a), \
+                numpy::aligned_array<type>(array), \
+                numpy::aligned_array<type>(Bc));
+
+        HANDLE_INTEGER_TYPES();
+#undef HANDLE
+    ...
+\end{cplusplus}
+
+These functions normally contain a lot of boiler-plate code: read the
+arguments, perform some sanity checks, perhaps a bit of initialisation, and
+then, the switch on the input type with the help of the
+\code{HANDLE\_INTEGER\_TYPES()}, \code{HANDLE\_FLOAT\_TYPE()}, and
+\code{HANDLE\_TYPES()} macros, which call the right specialisation of the
+template that does the actual work. In this example \code{erode} implements
+(binary) erosion:
+
+\begin{cplusplus}
+template<typename T>
+void erode(numpy::aligned_array<T> res,
+            numpy::aligned_array<T> array,
+            numpy::aligned_array<T> Bc) {
+    gil_release nogil;
+    const unsigned N = res.size();
+    typename numpy::aligned_array<T>::iterator iter = array.begin();
+    filter_iterator<T> filter(res.raw_array(), Bc.raw_array());
+    const unsigned N2 = filter.size();
+    T* rpos = res.data();
+
+    for (int i = 0;
+                i != N;
+                ++i, ++rpos, filter.iterate_both(iter)) {
+        for (int j = 0; j != N2; ++j) {
+            T arr_val = false;
+            filter.retrieve(iter, j, arr_val);
+            if (filter[j] && !arr_val) {
+                goto skip_this_one;
+            }
+        }
+        *rpos = true;
+        skip_this_one: continue;
+    }
+}
+\end{cplusplus}
+
+The template machinery is not that complicated and the functions using it are
+very simple and easy to read. The only downside is that there is some expansion
+of code size when the compiler instanciates the function for the several
+integer and floating point types. Given the small size of these functions, this
+is not a big issue.
+
+In the snippet above, you can see some other \cpp{} machinery:
+
+\begin{description}
+\item[\code{gil\_release}] This is a ``resource-acquisition is object
+initialisation'' (\textsc{raii})\footnote{\textsc{Raii} is a design pattern in
+\cpp{}, or other languages with scope linked deterministic object destruction,
+such as D, where a resource is represented by an object, whose constructor
+acquires it and whose destructor releases it. This guarantees that the object
+is correctly released even if the scope is left through an exception
+\citep{Stroustrup1994}.} object that release the Python global interpreter lock
+(\textsc{gil})\footnote{In the CPython interpreter, the most commonly used
+implementation of Python, there is a global lock for many Python related
+functionality, which limits parallelism.} in its constructor and gets it back
+in its destructor. Normally, the template function will release the
+\textsc{gil} after the Python-specific code is done.
+\item[\code{array}] This is a thin wrapper around \code{PyArrayObject} that
+knows its data type and has iterators which resemble the \cpp{} standard
+library.
+\item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it
+is useful to iterate over an image and use a centered filter around each pixel
+(it keeps track of all of the boundary conditions).
+\end{description}
+
+The inner loop is as direct an implementation of erosion as one would wish for:
+for each pixel in the image, look at its neighbours. If all are true, then set
+the corresponding output pixel to \textbf{true} (else, skip it as it has been
+initialised to zero).
+
+Most of the functions follow this architecture.
+
+\medskip
+
+Mahotas includes a full test suite. Every function has at least one
+corresponding test case. There are no known bugs.
+
+It has been available in the Python Package Index since April~2010 and has been
+downloaded over 5500~times. Mahotas is available as a FreeBSD package and is
+under evaluation for inclusion in the Debian distribution (which will cause it
+to be available as a native package for its derivatives, most famously Ubuntu).
+
+\subsection*{Acknowledgements}
+
+Mahotas includes code ported and incorporated from other projects. In
+particular, the \textsc{surf} implementation is a port from the code from
+\textit{dlib},\footnote{Dlib's webpage is at \url{http://dlib.net}.} a very
+good \cpp{} library by Davis King. I also gleaned some insight into the
+implementation of these features from Christopher Evan's OpenSURF library and
+its documentation \citep{evans2009}.\footnote{OpenSURF is available at
+\url{http://www.chrisevansdev.com/computer-vision-opensurf.html}, where several
+documents describe details of the implementation.} The image loading code,
+which interfaces with the FreeImage library, was written by Zachary Pincus and
+some of the support code was written by Peter J. Verveer for the
+\code{scipy.ndimage} project. All of these contributions were integrated while
+respecting the software licenses under which the original code had been
+released. Robert Webb, a summer student, worked with me on the local binary
+patterns implementation. Finally, I thank the several users who have reported
+bugs and submitted small fixes and participated on the project mailing list.
+
 \printbibliography

### 2012-07-04 paper.tex
MIN Update download counts

@@
 It has been available in the Python Package Index since April~2010 and has been
-downloaded over 5500~times. Mahotas is available as a FreeBSD package and is
-under evaluation for inclusion in the Debian distribution (which will cause it
-to be available as a native package for its derivatives, most famously Ubuntu).
+downloaded over 20,000~times. This does not include any downloads from other
+sources.
 
@@
 \url{http://www.chrisevansdev.com/computer-vision-opensurf.html}, where several
-documents describe details of the implementation.} The image loading code,
-which interfaces with the FreeImage library, was written by Zachary Pincus and
-some of the support code was written by Peter J. Verveer for the
-\code{scipy.ndimage} project. All of these contributions were integrated while
-respecting the software licenses under which the original code had been
-released. Robert Webb, a summer student, worked with me on the local binary
-patterns implementation. Finally, I thank the several users who have reported
-bugs and submitted small fixes and participated on the project mailing list.
+documents describe details of the implementation.} The code which interfaces
+with the FreeImage library, was written by Zachary Pincus and some of the
+support code was written by Peter J. Verveer for the \code{scipy.ndimage}
+project. All of these contributions were integrated while respecting the
+software licenses under which the original code had been released. Robert Webb,
+a summer student, worked with me on the local binary patterns implementation.
+Finally, I thank the several users who have reported bugs and submitted small
+fixes and participated on the project mailing list.
 

### 2012-07-05 paper.tex
TXT Start writing
I will switch to MIT License before releasing 1.0 and this paper

@@
 \section*{Abstract}
+Mahotas is a computer vision library for Python. It contains traditional image
+processing functionality such as filtering and morphological operations as well
+as more modern computer vision functions for feature computation, including
+local features.
+
+The interface is in Python, a dynamic programming language, which is very
+appropriate for fast development, but the algorithms are implemented in \cpp{}
+and are tuned for speed.
+
+Mahotas is available under a liberal open source license (MIT License) and is
+available from \url{http://github.com/luispedro/mahotas} or the Python Package
+Index (\url{http://pypi.python.org/pypi/mahotas}).
+
 \section{Introduction}
 
-Mahotas is a computer vision library for Python. It operates on numpy arrays,
+Mahotas is a computer vision library for the Python. It operates on numpy arrays,
 but its inner loops are implemented in \cpp{} for speed and ease of
@@
 
-Often, functions take an argument named \code{output} where the output will be
+Often, functions take an argument named \code{out} where the output will be
 stored. This argument is often much more restricted in type. In particular, it
@@
 
-The main reason that mahotas is in \cpp{} (and not in pure C) is to use templates.
-Almost \cpp{} functionality is split across 2~functions:
+The main reason that mahotas is implemented in \cpp{} (and not in C, which is
+the language of the Python interpreted) is to use templates. Almost \cpp{}
+functionality is split across 2~functions:
 
@@
 \item A \code{py\_function} which uses the Python C~API to get arguments and
-check them. This is almost always pure~C.
+check them.
 \item A template \code{function<dtype>} which works for the type \code{dtype}
@@
 
-The template machinery is not that complicated and the functions using it are
-very simple and easy to read. The only downside is that there is some expansion
-of code size when the compiler instanciates the function for the several
-integer and floating point types. Given the small size of these functions, this
-is not a big issue.
+The template machinery makes the functions that use it very simple and easy to
+read. The only downside is that there is some expansion of code size when the
+compiler instanciates the function for the several integer and floating point
+types. Given the small size of these functions, this is not a big issue.
 
@@
 in its destructor. Normally, the template function will release the
-\textsc{gil} after the Python-specific code is done.
+\textsc{gil} after the Python-specific code is done. This allows several
+mahotas functions to run concurrently.
 \item[\code{array}] This is a thin wrapper around \code{PyArrayObject} that
 knows its data type and has iterators which resemble the \cpp{} standard
-library.
+library. This makes the code type-safer.
 \item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it
@@
 
-\medskip
+\section{Discussion}
 
@@
 software licenses under which the original code had been released. Robert Webb,
-a summer student, worked with me on the local binary patterns implementation.
-Finally, I thank the several users who have reported bugs and submitted small
-fixes and participated on the project mailing list.
+a summer student at Carnegie Mellon University, worked with me on the initial
+local binary patterns implementation. Finally, I thank the several users who
+have reported bugs, submitted small fixes, and participated on the project
+mailing list.
+
+\textbf{Funding}: I was supported in my work by the Funda\c c\~{a}o para a
+Ci\^encia e Tecnologia (grants SFRH/BD/37535/2007 and PTDC/SAU-GMG/115652/2008)
+and by a grant from the Siebel Scholars Foundation.
 

### 2012-07-13 paper.tex
MIN Minor improvements

--- 2012-07-04 paper.tex
@@
 initialised to zero).
-
-Most of the functions follow this architecture.
 

### 2012-07-16 paper.tex
MIN A note about ml libraries

--- 2012-07-05 paper.tex
@@
 
+Mahotas does not include machine learning related functionality, such as
+$k$-means clustering or any classification. This is the result of an explicit
+design decision. Specialised machine learning packages for Python (e.g.,
+milk\footnote{http://github.com/luispedro/milk},
+scikit-learn\footnote{http://}) already exist. A good classification system can
+benefit both computer vision users and others. As these projects all use Numpy
+arrays as their datatypes, it is easy to use functionality from the different
+project seamlessly (including without any extra copying of data).
+
 Mahotas includes a full test suite. Every function has at least one
@@
 It has been available in the Python Package Index since April~2010 and has been
-downloaded over 20,000~times. This does not include any downloads from other
-sources.
+downloaded over 20,000~times\footnote{This does not include any downloads from
+other sources.}.
 

### 2012-08-08 paper.tex
TXT I just added wavelets to mahotas
ENH Less meaningless details, more meaningful ones

@@
 independently of each other (there is code sharing at the implementation
-level).
+level, but this is hidden from the user).
+
+The main functionality is grouped into the following categories:
 
 \begin{description}
+\item[\textsc{surf}] Speeded-up Robust Features.
+\item[Wavelet] Haar and Daubechies wavelets.
 \item[features] Some feature descriptors. In particular, Haralick texture
@@
 \item[watershed] seeded watershed.
-\item[thresholding] some thresholding methods.
 \item[polygon operations] convex hull, polygon drawing.
-\item[\textsc{surf}] Speeded-up Robust Features.
 \end{description}
@@
 When meaningful, a structuring element is used to define neighbourhoods or
-adjacency relationships (some functions which use this concept are: watershed,
-erosion, and median filtering). Generally, the default is to use a $3 \times 3$
-cross as the default if no structuring filter is given (the exception to this
-rule is the median filter, where the default is a $3 \times 3$ square).
+adjacency relationships (morphological functions, in particular, use this
+convention). Generally, the default is to use a $3 \times 3$ cross as the
+default if no structuring filter is given (the exception to this rule is the
+median filter, where the default is a $3 \times 3$ square).
 
@@
 
-These functions normally contain a lot of boiler-plate code: read the
-arguments, perform some sanity checks, perhaps a bit of initialisation, and
-then, the switch on the input type with the help of the
-\code{HANDLE\_INTEGER\_TYPES()}, \code{HANDLE\_FLOAT\_TYPE()}, and
-\code{HANDLE\_TYPES()} macros, which call the right specialisation of the
-template that does the actual work. In this example \code{erode} implements
-(binary) erosion:
+This functions retrieves the arguments, performs some sanity checks, performs a
+bit of initialisation, and finally, switches in the input type with the help of
+the \code{HANDLE\_INTEGER\_TYPES()} macro, which call the right specialisation
+of the template that does the actual work. In this example \code{erode}
+implements (binary) erosion:
 
@@
 
-Most of the functions follow this architecture.
-
 \section{Discussion}
@@
 
-Mahotas includes a full test suite. Every function has at least one
-corresponding test case. There are no known bugs.
+Python is an ideal language for fast development of both applications and
+scientific software. With a fast library of computer vision and image
+processing functions (implemented in \cpp{}, as the standard Python interpreter
+is too slow for a direct Python implementation).
 
-It has been available in the Python Package Index since April~2010 and has been
-downloaded over 20,000~times\footnote{This does not include any downloads from
-other sources.}.
+Mahotas has been available in the Python Package Index since April~2010 and has
+been downloaded over 20,000~times. This does not include any downloads from
+other sources. Mahotas includes a full test suite (version 1.0 has 100\% test
+coverage). There are no known bugs.
 

### 2012-08-11 paper.tex
MIN A note about version numbers

@@
 
-\subsection{Functionality and Interface}
+Version 1.0 has been released recently and this is now a mature, well-tested
+package (the first version was made available over 4~years ago). \footnote{Note
+for reviewers: the version currently available implements all functionality
+described in this manuscript. I will release it as version 1.0 when this
+manuscript is accepted. Naturally, any bugs that are reported in the
+meanwhile will be addressed.}
+
+\section{Functionality and Interface}
 
@@
 
-\subsection{Implementation}
+\section{Implementation}
 

### 2012-08-20 paper.tex
MIN A note about numpy
This is one thing to emphacise: using Python, we build on other
structures, instead of defining our own.

@@
 
-Mahotas is a computer vision library for the Python. It operates on numpy arrays,
-but its inner loops are implemented in \cpp{} for speed and ease of
-implementation.
+Mahotas is a computer vision library for the Python. It operates on numpy
+arrays~\cite{numpy}. Therefore, it uses all the infrastructure built by that
+project for storing information. However, its inner loops are implemented in
+\cpp{} for speed and ease of implementation.
 

### 2012-08-21 paper.tex
ENH Add references to other systems

@@
 Mahotas is a computer vision library for the Python. It operates on numpy
-arrays~\cite{numpy}. Therefore, it uses all the infrastructure built by that
-project for storing information. However, its inner loops are implemented in
-\cpp{} for speed and ease of implementation.
+arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
+by that project for storing information.
+
+Python is a natural ``glue'' language: it is easy to use state-of-the-art
+libraries written in multiple languages~\citep{10.1109/MCSE.2007.58}. Mahotas
+itself is a mix of high-level Python and low-level \cpp{}. This achieves a good
+balance between speed and ease of implementation.
 
@@
 \begin{description}
-\item[\textsc{surf}] Speeded-up Robust Features.
+\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}.
 \item[Wavelet] Haar and Daubechies wavelets.
@@
 milk\footnote{http://github.com/luispedro/milk},
-scikit-learn\footnote{http://}) already exist. A good classification system can
-benefit both computer vision users and others. As these projects all use Numpy
-arrays as their datatypes, it is easy to use functionality from the different
-project seamlessly (including without any extra copying of data).
+scikit-learn\footnote{http://scikit-learn}) already
+exist~\citep{Pedregosa:2011:SML:2078183.2078195}. A good classification (or
+other) system can benefit both computer vision users and others. As these
+projects all use Numpy arrays as their data types, it is easy to use
+functionality from the different project seamlessly (no copying of data is
+necessary).
 

### 2012-08-22 paper.tex
MIN This bears saying
ENH More details

@@
 \title{Mahotas: Open source software for scriptable computer vision}
-\author{Luis Pedro Coelho}
+\author{Luis Pedro Coelho\\
+Lane Center for Computational Biology, Carnegie Mellon University, Pittsburgh, USA\\
+Instituto de Medicina Molecular, Lisboa, Portugal}
 
@@
 arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
-by that project for storing information.
+by that project for storing information. In particular, unlike libraries
+written in the C~Language or in Java, Mahotas does not need to define a data
+structure, but uses the numpy array. Many basic manipulation functionality that
+would otherwise be part of a computer vision library are handled by numpy, for
+example computing averages and other simple statistics, handling multi-channel
+images, converting between types (integer and floating point images are
+supported by mahotas, whenever it is meaningful). For the user, this has the
+additional advantage that they do not need to learn yet another set of
+functions.
+
+On the other end, by integrating into the Python numeric ecosystem, users can
+use other packages in a seamless way. In particular, mahotas does not implement
+any machine learning functionality, but rather advises the user to use another,
+specialised package, such as scikits-learn or milk.
 
@@
 
-Version 1.0 has been released recently and this is now a mature, well-tested
-package (the first version was made available over 4~years ago). \footnote{Note
-for reviewers: the version currently available implements all functionality
-described in this manuscript. I will release it as version 1.0 when this
-manuscript is accepted. Naturally, any bugs that are reported in the
-meanwhile will be addressed.}
+Version 1.0 of mahotas has been released recently and this is now a mature,
+well-tested package (the first version was made available over 4~years ago).
+\footnote{Note for reviewers: the version currently available implements all
+functionality described in this manuscript. I will release it as version 1.0
+when this manuscript is accepted to coincide with publication. Naturally, any
+bugs that are found and reported in the meanwhile will be addressed.} Mahotas
+runs and is used on different versions of Unix (including Linux, SunOS, and
+FreeBSD), Mac OS X, and Windows.
 
@@
 \begin{description}
-\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}.
-\item[Wavelet] Haar and Daubechies wavelets.
-\item[features] Some feature descriptors. In particular, Haralick texture
+\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This is
+a collection of local features.
+\item[features] Global feature descriptors. In particular, Haralick texture
 features, Zernike moments, local binary patterns, threshold adjacency
 statistics.
+\item[Wavelet] Haar and Daubechies wavelets. Forward and reverse transforms.
 \item[morphological functions] Erosion and dilation, as well as some more
-complex operations.
-\item[watershed] seeded watershed.
+complex operations built on these. There are both binary and grayscale
+implementations of these operators.
+\item[watershed] seeded watershed and distance map transforms.
+\item[filtering] Gaussian filtering and general convolutions.
 \item[polygon operations] convex hull, polygon drawing.
@@
 
-There are a few interface conventions which apply to many functions.
+In general, functions operate on all data types (except when it not well
+defined). This is performed without any extra memory copies. Mahotas is heavily
+optimised for both speed and memory usage.
 
-When meaningful, a structuring element is used to define neighbourhoods or
-adjacency relationships (morphological functions, in particular, use this
-convention). Generally, the default is to use a $3 \times 3$ cross as the
+There are a few interface conventions which apply to many functions. When
+meaningful, a structuring element is used to define neighbourhoods or adjacency
+relationships (morphological functions, in particular, use this convention).
+Generally, the default is to use a $3 \times 3$ cross as the
 default if no structuring filter is given (the exception to this rule is the
@@
 
-Mahotas is written in \cpp, but almost always, the user calls a Python function
-which checks types and then calls the internal function. This is slightly
-slower, but it is easier to develop this way.
+Mahotas is written in \cpp, but this is completely hidden from the user as
+there are hand-written Python wrappers for all functions (automatically
+generated wrappers inevitably lead to worse error messages and are less
+flexible).
 
@@
 
+Mahotas uses the standard Python build machinery and distribution channels.
+Building and installing from source code is done using
+\begin{verbatim}
+python setup.py install
+\end{verbatim}
+Alternatively, Python based package managers (such as \texttt{easy\_install} or
+\texttt{pip}) can be used (mahotas works well with these systems).
+
 \section{Discussion}

### 2012-08-22 paper.tex
TXT+FIG Add example usage

@@
 one).
+
+\subsection{Example of Use}
+
+This is a simple example of mahotas usage. Code for this and other examples is
+present in the mahotas source distribution under the \texttt{demos/} directory.
+
+In this tutorial, we load an image and find SURF interest point and descriptors
+on it.
+
+We start by importing the necessary packages, including numpy and mahotas. We
+also use \textit{milk}, to demonstrate how the mahotas output can integrate
+with a machine learning package.
+
+\begin{python}
+import numpy as np
+import mahotas
+from mahotas.features import surf
+import milk
+\end{python}
+
+The first step is to load the image and convert to 8~bit numbers.
+\begin{python}
+f = mahotas.imread('luispedro.jpg', as_grey=True)
+f = f.astype(np.uint8)
+\end{python}
+
+We can now compute SURF interest points and descriptors.
+\begin{python}
+spoints = surf.surf(f, 4, 6, 2)
+\end{python}
+
+Using numpy operations and milk, we can select only the descriptors and cluster
+them into a group of five:
+
+\begin{python}
+descrs = spoints[:,6:]
+
+values, _ = milk.kmeans(descrs, 5)
+\end{python}
+
+Finally, we can show the 
+\begin{python}
+colors = np.array(
+    [ 255,  25,   1],
+    [203,  77,  37],
+    [151, 129,  56],
+    [ 99, 181,  52],
+    [ 47, 233,   5]])
+f2 = surf.show_surf(f, spoints[:64], values, colors)
+\end{python}
+
+Using matplotlib~\citep{10.1109/MCSE.2007.55}, we finally display the image as
+Figure~\ref{fig:surf}.
+
+\begin{python}
+from matplotlib import pyplot as plt
+plt.subplot(1,2,1)
+plt.imshow(f)
+plt.subplot(1,2,2)
+plt.imshow(f2)
+\end{python}
+
+The easy interaction with matplotlib is another way in which we benefit from
+the numpy-based ecosystem.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=.8\textwidth]{surf-tutorial}
+\end{center}
+\caption{Example of Usage. On the left, the original image is shown, while on
+the right SURF detections are represented as rectangles of different colours.}
+\label{fig:surf}
+\end{figure}
 

### 2012-08-22 figures/surf-tutorial.py

@@ -1,0 +1,26 @@
+import numpy as np
+import mahotas
+from mahotas.features import surf
+from os import path
+from pylab import *
+import milk
+
+f = mahotas.imread('luispedro.jpg', as_grey=True)
+f = f.astype(np.uint8)
+spoints = surf.surf(f, 4, 6, 2)
+descrs = spoints[:,6:]
+
+values, _ = milk.kmeans(descrs, 5)
+colors = np.array([(255-52*i,25+52*i,37**i % 101) for i in xrange(5)])
+
+f2 = surf.show_surf(f, spoints[:64], values, colors)
+
+subplot(1,2,1)
+imshow(np.dstack([f,f,f]))
+xticks([])
+yticks([])
+subplot(1,2,2)
+imshow(f2)
+xticks([])
+yticks([])
+savefig('surf-tutorial.png')

### 2012-08-22 paper.tex
TXT More explicit comparison with alternatives

@@
 by that project for storing information. In particular, unlike libraries
-written in the C~Language or in Java, Mahotas does not need to define a data
-structure, but uses the numpy array. Many basic manipulation functionality that
-would otherwise be part of a computer vision library are handled by numpy, for
-example computing averages and other simple statistics, handling multi-channel
-images, converting between types (integer and floating point images are
-supported by mahotas, whenever it is meaningful). For the user, this has the
-additional advantage that they do not need to learn yet another set of
-functions.
+written in the C~Language or in Java~\citep{Marcel:2010:TMP:1873951.1874254},
+Mahotas does not need to define a data structure, but uses the numpy array
+structure. Many basic manipulation functionality that would otherwise be part
+of a computer vision library are handled by numpy, for example computing
+averages and other simple statistics, handling multi-channel images, converting
+between types (integer and floating point images are supported by mahotas,
+whenever it is meaningful). For the user, this has the additional advantage
+that they do not need to learn yet another set of functions.
 
@@
 Python is an ideal language for fast development of both applications and
-scientific software. With a fast library of computer vision and image
-processing functions (implemented in \cpp{}, as the standard Python interpreter
-is too slow for a direct Python implementation).
+scientific software. For this platform, Mahotas is a fast library of computer
+vision and image processing functions. It is implemented in \cpp{}, as the
+standard Python interpreter is too slow for a direct Python implementation.
+However, all of the Python interface code is hand-written, as opposed to using
+automatic interface generators like Swig~\cite{Beazley2003599}. This is more
+work, but the end result is of much higher quality, especially when it comes to
+giving useful error messages (e.g., when a type mismatch occurs, an automatic
+system will often be forced to resort to a generic message as it does not have
+any knowledge of what the arguments mean besides their automatically inferred
+types).
 

### 2012-08-24 paper.tex
TXT Mahotas works for Python 3 now
BIB Reference distance map algorithm
I think this is one of the few that is young enough to not be just part
of the culture (along with SURF, of course).

@@
 
-Mahotas is a computer vision library for the Python. It operates on numpy
+Mahotas is a computer vision library for the Python Programming Language
+(versions 2.5 and up, including version 3). It operates on numpy
 arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
@@
 implementations of these operators.
-\item[watershed] seeded watershed and distance map transforms.
+\item[watershed] seeded watershed and distance map
+transforms~\citep{felzenszwalb}.
 \item[filtering] Gaussian filtering and general convolutions.

### 2012-08-28 paper.tex
ENH Add information fields

@@
 
+\textbf{Keywords:} computer vision, image processing.
+
 \section{Introduction}
@@
 
-\section{Functionality and Interface}
+\section{Implementation and Architecture}
+
+\subsection{Interface}
 
@@
 statistics.
-\item[Wavelet] Haar and Daubechies wavelets. Forward and reverse transforms.
+\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
 \item[morphological functions] Erosion and dilation, as well as some more
@@
 relationships (morphological functions, in particular, use this convention).
-Generally, the default is to use a $3 \times 3$ cross as the
-default if no structuring filter is given (the exception to this rule is the
-median filter, where the default is a $3 \times 3$ square).
+Generally, the default is to use a $3 \times 3$ cross as the default if no
+structuring filter is given (the exception to this rule is the median filter,
+where the default is a $3 \times 3$ square).
 
@@
 
-\section{Implementation}
+\subsection{Implementation}
 
@@
 
+\subsection{Quality Control}
+
+Mahotas includes a complete automated suite of unit tests. These all
+functionality and include several regression tests. There are no known bugs in
+version~1.0. In fact, no releases have ever been performed with known bugs.
+Naturally, bugs were, occasionally, discovered in released versions, but
+corrected before the next release.
+
+The development is completely open-source and development versions are
+available. Many users have submitted bug reports and fixes.
+
+\section{Availability}
+
+\textbf{Operating system}\\
+Mahotas runs and is used on different versions of Unix (including Linux, SunOS,
+and FreeBSD), Mac OS X, and Windows.
+
+\textbf{Programming language}\\
+Mahotas works in Python (minimal version is~2.5 and works in all more recent
+versions, including in the Python~3 series).
+
+\textbf{Additional system requirements}\\
+None.
+
+\textbf{Dependencies}\\
+It requires numpy to be present and installed.
+
+\textbf{List of contributors}\\
+Luis Pedro Coelho, Zachary Pincus, Peter J. Verveer, Davis King, Robert Webb
+(Carnegie Mellon University), Matthew Goodman, K.-Michael Aye (University of
+Bern), Rita Simões, Joe Kington (University of Wisconsin Dept. of Geoscience),
+Christoph Gohlke (University of California, Irvine), and Sandro Knauss.
+
+\subsubsection{Software location}
+
+\textbf{Code repository}\\
+\textit{Name}: Github\\
+\textit{Identifier}: https://github.com/luispedro/mahotas\\
+\textit{Licence}: MIT\\
+\textit{Date published}: Since 2009
+
 \section{Discussion}
@@
 
+\section{Reuse Potential}
+
+Originally, this code was developed in the context of cellular image analysis.
+However, none of the functionality is specific to this context and many
+computer vision pipelines can make use of it.
+
 \subsection*{Acknowledgements}

### 2012-08-29 paper.tex
TXT Add references to papers using mahotas

@@
 
+This package (and earlier versions of it) have been used by
+myself~\citep{Coelho2009,Coelho2010a} and close collaborators in several
+publications~\citep{omerosearcher}. Other groups have used in published work,
+both in cell image analysis~\citep{CYTO:CYTO22034} and in other
+areas~\citep{springerlink:10.1007/978-3-642-32335-5_2}.
+
 \subsection*{Acknowledgements}

### 2012-08-29 paper.tex
TXT Typo. Small note

@@
 The main reason that mahotas is implemented in \cpp{} (and not in C, which is
-the language of the Python interpreted) is to use templates. Almost \cpp{}
+the language of the Python interpreter) is to use templates. Almost \cpp{}
 functionality is split across 2~functions:
@@
 knows its data type and has iterators which resemble the \cpp{} standard
-library. This makes the code type-safer.
+library. This makes the code type-safer. This is also a \textsc{raii} object in
+terms of managing Python reference counts.
 \item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it

### 2012-08-29 paper.tex
BIB Add affiliations to contributors

@@
 \textbf{List of contributors}\\
-Luis Pedro Coelho, Zachary Pincus, Peter J. Verveer, Davis King, Robert Webb
-(Carnegie Mellon University), Matthew Goodman, K.-Michael Aye (University of
-Bern), Rita Simões, Joe Kington (University of Wisconsin Dept. of Geoscience),
-Christoph Gohlke (University of California, Irvine), and Sandro Knauss.
+Luis Pedro Coelho (Carnegie Mellon University and Instituto de Medicina
+Molecular), Zachary Pincus (Stanford University), Peter J. Verveer (European
+Molecular Biology Laboratory), Davis King (Northrop Grumman ES), Robert Webb
+(Carnegie Mellon University), Matthew Goodman (University of Texas at Austin),
+K.-Michael Aye (University of Bern), Rita Sim\~{o}es (University of Twente),
+Joe Kington (University of Wisconsin Dept.  of Geoscience), Christoph Gohlke
+(University of California, Irvine), and Sandro Knauss (University of Bremen).
 

### 2012-08-29 paper.tex
TXT Machine Learning Packages

@@
 
-The interface is a procedural interface, with most functions working
-independently of each other (there is code sharing at the implementation
+The interface is a procedural interface, with no global state. All functions
+work independently of each other (there is code sharing at the implementation
 level, but this is hidden from the user).
@@
 \begin{description}
-\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This is
-a collection of local features.
+\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This
+includes both keypoint detection and descriptor computation.
 \item[features] Global feature descriptors. In particular, Haralick texture
-features, Zernike moments, local binary patterns, threshold adjacency
-statistics.
+features, Zernike moments, local binary patterns, and threshold adjacency
+statistics (both the original~\citep{Hamilton2007} and the parameter-free
+versions~\citep{Coelho2010}).
 \item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
@@
 
-In general, functions operate on all data types (except when it not well
-defined). This is performed without any extra memory copies. Mahotas is heavily
-optimised for both speed and memory usage.
+In general, functions operate on all data types (in some cases, the operation
+is only meaningful for floating point arrays, for example). This is performed
+without any extra memory copies. Mahotas is heavily optimised for both speed
+and memory usage (it can be used with very large arrays).
 
@@
 
-Often, functions take an argument named \code{out} where the output will be
-stored. This argument is often much more restricted in type. In particular, it
-must often be a contiguous array.\footnote{Numpy supports non-contiguous
-arrays, which are most often slices into other, larger, contiguous arrays
-(e.g., given a $128 \times 128$ contiguous array, one can build a $64 \times
-128$ non-contiguous array by taking every other row).} Since this is a
-performance feature, it is natural that the interface is less flexible
-(accessing a contiguous array is much more efficient than a non-contiguous
-one).
+When a new image is to be returned, functions take an argument named \code{out}
+where the output will be stored. This argument is often much more restricted in
+type. In particular, it must often be a contiguous array.\footnote{Numpy
+supports non-contiguous arrays, which are most often slices into other, larger,
+contiguous arrays (e.g., given a $128 \times 128$ contiguous array, one can
+build a $64 \times 128$ non-contiguous array by taking every other row).} Since
+this is a performance feature (its purpose is to avoid extra memory
+allocation), it is natural that the interface is less flexible (accessing a
+contiguous array is much more efficient than a non-contiguous one).
 
@@
 present in the mahotas source distribution under the \texttt{demos/} directory.
-
-In this tutorial, we load an image and find SURF interest point and descriptors
+In this example, we load an image and find SURF interest point and descriptors
 on it.
@@
 
-The first step is to load the image and convert to 8~bit numbers.
+The first step is to load the image and convert to 8~bit numbers. In this
+case, the conversion is done using standard numpy methods, namely
+\code{astype}:
+
 \begin{python}
@@
 descrs = spoints[:,6:]
-
 values, _ = milk.kmeans(descrs, 5)
@@
 
-Finally, we can show the 
+Finally, we can show the points in colour.
 \begin{python}
@@
 
-Using matplotlib~\citep{10.1109/MCSE.2007.55}, we finally display the image as
-Figure~\ref{fig:surf}.
+The \texttt{show\_surf} only builds the image as a multi-channel (one for each
+colour) image. Using matplotlib~\citep{10.1109/MCSE.2007.55}, we finally
+display the image as Figure~\ref{fig:surf}.
 
@@
 
-Mahotas is written in \cpp, but this is completely hidden from the user as
-there are hand-written Python wrappers for all functions (automatically
+Mahotas is mostly written in \cpp, but this is completely hidden from the user
+as there are hand-written Python wrappers for all functions (automatically
 generated wrappers inevitably lead to worse error messages and are less
@@
 
-Mahotas uses the standard Python build machinery and distribution channels.
-Building and installing from source code is done using
+In keeping with the philosophy of blending in with the ecosystem, Mahotas uses
+the standard Python build machinery and distribution channels. Building and
+installing from source code is done using
 \begin{verbatim}
@@
 K.-Michael Aye (University of Bern), Rita Sim\~{o}es (University of Twente),
-Joe Kington (University of Wisconsin Dept.  of Geoscience), Christoph Gohlke
-(University of California, Irvine), and Sandro Knauss (University of Bremen).
+Joe Kington (University of Wisconsin), Christoph Gohlke (University of
+California, Irvine), and Sandro Knauss (University of Bremen).
 
@@
 $k$-means clustering or any classification. This is the result of an explicit
-design decision. Specialised machine learning packages for Python (e.g.,
-milk\footnote{http://github.com/luispedro/milk},
-scikit-learn\footnote{http://scikit-learn}) already
-exist~\citep{Pedregosa:2011:SML:2078183.2078195}. A good classification (or
-other) system can benefit both computer vision users and others. As these
-projects all use Numpy arrays as their data types, it is easy to use
-functionality from the different project seamlessly (no copying of data is
-necessary).
+design decision. Specialised machine learning packages for Python already
+exist~\citep{Pedregosa:2011:SML:2078183.2078195,springerlink:10.1007/978-3-540-30116-5_58,Schaul:2010:PYB:1756006.1756030,Sonnenburg:2010:SML:1756006.1859911}.
+A good classification (or other) system can benefit both computer vision users
+and others. As these projects all use Numpy arrays as their data types, it is
+easy to use functionality from the different project seamlessly (no copying of
+data is necessary).
 

### 2012-08-30 timethings.py
TAB Add timings table

@@ -1,0 +1,45 @@
+import numpy as np
+import timeit
+import mahotas
+luispedro_image = 'mahotas/demos/data/luispedro.jpg'
+f = mahotas.imread(luispedro_image, as_grey=True)
+markers = np.zeros_like(f)
+markers[100,100] = 1
+markers[200,200] = 2
+f = f.astype(int)
+markers = markers.astype(int)
+otsu = mahotas.otsu(f.astype(np.uint8))
+fbin = f > otsu
+Bc = np.eye(3)
+Bc = Bc.astype(bool)
+
+pre ='''
+import numpy as np
+import mahotas
+import pymorph
+import timethings
+fbin = timethings.fbin
+f = timethings.f
+Bc = timethings.Bc
+markers = timethings.markers
+'''
+
+def t(s):
+    return timeit.timeit(s, setup=pre, number=24)
+if __name__ == '__main__':
+    base = t('np.max(f)')
+    for name,statement, alt in [
+                ('erode', 'mahotas.erode(fbin, Bc)', 'pymorph.erode(fbin, Bc)'),
+                ('dilate', 'mahotas.dilate(fbin, Bc)', 'pymorph.dilate(fbin, Bc)'),
+                ('open', 'mahotas.open(fbin, Bc)', 'pymorph.open(fbin, Bc)'),
+                ('center mass', 'mahotas.center_of_mass(f)', None),
+                ('cwatershed', 'mahotas.cwatershed(f, markers)', 'pymorph.cwatershed(f, markers)'),
+                ('daubechies', 'mahotas.daubechies(f, "D4")', None),
+                ('haralick', 'mahotas.features.haralick(f)', None),
+                ]:
+        time = '%.2f' % (t(statement)/base)
+        alttime = 'NA'
+        if alt is not None:
+            alttime = '%.2f' % (t(alt)/base)
+        print r"%-12s& %8s & %8s \\" % (name, time, alttime)
+

### 2012-08-30 paper.tex
TXT Add timings. Fix typos. Reorg

@@
 \usepackage{hyperref}
+\usepackage{booktabs}
 \usepackage{tikz}
@@
 as more modern computer vision functions for feature computation, including
-local features.
+interest point detection and local descriptors.
 
@@
 arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
-by that project for storing information. In particular, unlike libraries
-written in the C~Language or in Java~\citep{Marcel:2010:TMP:1873951.1874254},
-Mahotas does not need to define a data structure, but uses the numpy array
-structure. Many basic manipulation functionality that would otherwise be part
-of a computer vision library are handled by numpy, for example computing
-averages and other simple statistics, handling multi-channel images, converting
-between types (integer and floating point images are supported by mahotas,
-whenever it is meaningful). For the user, this has the additional advantage
-that they do not need to learn yet another set of functions.
+by that project for storing information and performing basic manipulations and
+computations. In particular, unlike libraries written in the C~Language or in
+Java~\citep{Marcel:2010:TMP:1873951.1874254}, Mahotas does not need to define a
+data structure, but uses the numpy array structure. Many basic manipulation
+functionality that would otherwise be part of a computer vision library are
+handled by numpy, for example computing averages and other simple statistics,
+handling multi-channel images, converting between types (integer and floating
+point images are supported by mahotas, whenever it is meaningful). For the
+user, this has the additional advantage that they do not need to learn yet
+another set of functions.
 
-On the other end, by integrating into the Python numeric ecosystem, users can
-use other packages in a seamless way. In particular, mahotas does not implement
-any machine learning functionality, but rather advises the user to use another,
+Additionally, by integrating into the Python numeric ecosystem, users can use
+other packages in a seamless way. In particular, mahotas does not implement any
+machine learning functionality, but rather advises the user to use another,
 specialised package, such as scikits-learn or milk.
@@
 versions~\citep{Coelho2010}).
-\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
+\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
 \item[morphological functions] Erosion and dilation, as well as some more
@@
 
-In general, functions operate on all data types (in some cases, the operation
-is only meaningful for floating point arrays, for example). This is performed
-without any extra memory copies. Mahotas is heavily optimised for both speed
-and memory usage (it can be used with very large arrays).
+In general, functions operate on all data types. This is performed without any
+extra memory copies. Mahotas is heavily optimised for both speed and memory
+usage (it can be used with very large arrays).
 
@@
     PyArrayObject* Bc;
-    if (!PyArg_ParseTuple(args,"OO", &array, &Bc)) {
+    PyArrayObject* output;
+    if (!PyArg_ParseTuple(args, "OOO", &array, &Bc, &output) ||
+        !numpy::are_arrays(array, Bc, output) || !numpy::same_shape(array, output) ||
+        !PyArray_EquivTypenums(PyArray_TYPE(array), PyArray_TYPE(Bc)) ||
+        !PyArray_EquivTypenums(PyArray_TYPE(array), PyArray_TYPE(output)) ||
+        PyArray_NDIM(array) != PyArray_NDIM(Bc)
+    ) {
+        PyErr_SetString(PyExc_RuntimeError, TypeErrorMsg);
         return NULL;
     }
-    PyArrayObject* res_a = (PyArrayObject*)PyArray_SimpleNew(
-                                array->nd,
-                                array->dimensions,
-                                PyArray_TYPE(array));
-    if (!res_a) return NULL;
-    PyArray_FILLWBYTE(res_a, 0);
-    switch(PyArray_TYPE(array)) {
+    holdref r_o(output);
+
 #define HANDLE(type) \
-    erode<type>(numpy::aligned_array<type>(res_a), \
+    erode<type>(numpy::aligned_array<type>(output), \
                 numpy::aligned_array<type>(array), \
                 numpy::aligned_array<type>(Bc));
-
-        HANDLE_INTEGER_TYPES();
+    SAFE_SWITCH_ON_INTEGER_TYPES_OF(array, true);
 #undef HANDLE
@@
 bit of initialisation, and finally, switches in the input type with the help of
-the \code{HANDLE\_INTEGER\_TYPES()} macro, which call the right specialisation
-of the template that does the actual work. In this example \code{erode}
-implements (binary) erosion:
+the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES()} macro, which call the right
+specialisation of the template that does the actual work. In this example
+\code{erode} implements (binary) erosion:
 
@@
     gil_release nogil;
-    const unsigned N = res.size();
+    const int N = res.size();
     typename numpy::aligned_array<T>::iterator iter = array.begin();
-    filter_iterator<T> filter(res.raw_array(), Bc.raw_array());
-    const unsigned N2 = filter.size();
+    filter_iterator<T> filter(array.raw_array(), Bc.raw_array(), EXTEND_NEAREST, is_bool(T()));
+    const int N2 = filter.size();
     T* rpos = res.data();
@@
                 ++i, ++rpos, filter.iterate_both(iter)) {
+        T value = std::numeric_limits<T>::max();
         for (int j = 0; j != N2; ++j) {
-            T arr_val = false;
+            T arr_val = T();
             filter.retrieve(iter, j, arr_val);
-            if (filter[j] && !arr_val) {
-                goto skip_this_one;
+            value = std::min<T>(value, erode_sub(arr_val, filter[j]));
         }
-        }
-        *rpos = true;
-        skip_this_one: continue;
+        *rpos = value;
     }
@@
 read. The only downside is that there is some expansion of code size when the
-compiler instanciates the function for the several integer and floating point
+compiler instantiates the function for the several integer and floating point
 types. Given the small size of these functions, this is not a big issue.
@@
 The inner loop is as direct an implementation of erosion as one would wish for:
-for each pixel in the image, look at its neighbours. If all are true, then set
-the corresponding output pixel to \textbf{true} (else, skip it as it has been
-initialised to zero).
+for each pixel in the image, look at its neighbours, subtract the filter value,
+and compute the minimum of this operation.
+
+\subsection{Efficiency}
+
+\begin{table}
+\centering
+\begin{tabular}{lrr}
+\toprule
+Operation & mahotas & pymorphy\\
+\midrule
+erode       &     9.90 &    13.06 \\
+dilate      &    10.48 &     8.17 \\
+open        &    20.26 &    21.05 \\
+cwatershed  &   181.47 & 50211.56 \\
+center mass &     6.39 &       NA \\
+daubechies  &    16.92 &       NA \\
+haralick    &   276.90 &       NA \\
+\bottomrule
+\end{tabular}
+\caption{Efficiency Results for mahotas and pymorph. Shown are values as
+multiples of the time that \code{numpy.max(f)} takes.}
+\label{tab:efficiency}
+\end{table}
+
+Table~\ref{tab:efficiency} shows timings for different operations. These were
+normalized to multiples of the time it takes to go over the image and find its
+maximum pixel value (this was done using \code{numpy.max(image}). The
+measurements shown were obtained on an Intel 64~bit system, running Ubuntu
+Linux. However, due to the normalization, measurements obtained on another
+system (Intel 32~bits running Mac OS) were qualitatively similar (data not
+shown).
+
+The comparison is against Pymorph~\citep{DougLotu:03}, which is a pure Python
+implementation of some of the same functions. We can see that simple
+morphological operations (erosion, dilation, opening) take a similar amount of
+time in mahotas and pymorph (pymorph intelligently uses numpy operations for
+these). However, some more complex operations such as watershed are much faster
+in mahotas.
+
+\bigskip
 
@@
 Mahotas runs and is used on different versions of Unix (including Linux, SunOS,
-and FreeBSD), Mac OS X, and Windows.
+and FreeBSD), Mac OS X, and Windows.\footnote{Christoph Gohlke has been
+instrumental in providing Windows packages as well as several fixes for that
+platform.}
 
@@
 
+\section{Reuse Potential}
+
+Originally, this code was developed in the context of cellular image analysis.
+However, none of the functionality is specific to this context and many
+computer vision pipelines can make use of it.
+
+This package (and earlier versions of it) have been used by
+myself~\citep{Coelho2009,Coelho2010a} and close collaborators in several
+publications~\citep{omerosearcher}. Other groups have used in published work,
+both in cell image analysis~\citep{CYTO:CYTO22034} and in other
+areas~\citep{springerlink:10.1007/978-3-642-32335-5_2}.
+
 \section{Discussion}
+
+Python is an excellent language for scientific programming because of the
+inherent properties of the language and because of the infrastructure that has
+been built around the numpy project. Mahotas works in this environment to
+provide the user with image analysis and computer vision functionality.
 
@@
 been downloaded over 20,000~times. This does not include any downloads from
-other sources. Mahotas includes a full test suite (version 1.0 has 100\% test
-coverage). There are no known bugs.
-
-\section{Reuse Potential}
-
-Originally, this code was developed in the context of cellular image analysis.
-However, none of the functionality is specific to this context and many
-computer vision pipelines can make use of it.
-
-This package (and earlier versions of it) have been used by
-myself~\citep{Coelho2009,Coelho2010a} and close collaborators in several
-publications~\citep{omerosearcher}. Other groups have used in published work,
-both in cell image analysis~\citep{CYTO:CYTO22034} and in other
-areas~\citep{springerlink:10.1007/978-3-642-32335-5_2}.
+other sources. Mahotas includes a full test suite. There are no known bugs.
 

### 2012-08-30 timethings.py
ENH More benchmarking

@@ -1,7 +1,10 @@
+import skimage.morphology
+import skimage.filter
+import skimage.feature
 import numpy as np
 import timeit
 import mahotas
-luispedro_image = 'mahotas/demos/data/luispedro.jpg'
+luispedro_image = '../mahotas/mahotas/demos/data/luispedro.jpg'
 f = mahotas.imread(luispedro_image, as_grey=True)
 markers = np.zeros_like(f)
 markers[100,100] = 1
@@ -10,36 +13,82 @@ f = f.astype(int)
 markers = markers.astype(int)
 otsu = mahotas.otsu(f.astype(np.uint8))
 fbin = f > otsu
+fbin8 = fbin.astype(np.uint8)
 Bc = np.eye(3)
 Bc = Bc.astype(bool)
+Bc8 = Bc.astype(np.uint8)
 
 pre ='''
+import skimage.filter
+import skimage.morphology
+import skimage.feature
 import numpy as np
 import mahotas
 import pymorph
 import timethings
-fbin = timethings.fbin
 f = timethings.f
+fbin = timethings.fbin
+fbin8 = timethings.fbin8
+f64 = f.astype(np.float64)
 Bc = timethings.Bc
+Bc8 = timethings.Bc8
 markers = timethings.markers
 '''
 
 def t(s):
     return timeit.timeit(s, setup=pre, number=24)
+
+tests = [
+    ('erode', [
+        'mahotas.erode(fbin, Bc)',
+        'pymorph.erode(fbin, Bc)',
+        'skimage.morphology.opening(fbin8, Bc8)',
+        ]),
+    ('dilate', [
+        'mahotas.dilate(fbin, Bc)',
+        'pymorph.dilate(fbin, Bc)',
+        'skimage.morphology.dilation(fbin8, Bc8)',
+        ]),
+    ('open', [
+        'mahotas.open(fbin, Bc)',
+        'pymorph.open(fbin, Bc)',
+        'skimage.morphology.opening(fbin8, Bc8)',
+        ]),
+    ('center mass', [
+        'mahotas.center_of_mass(f)',
+        None,
+        None,
+        ]),
+    ('sobel', [
+        'mahotas.sobel(f)',
+        None,
+        'skimage.filter.sobel(f64)',
+        ]),
+    ('cwatershed', [
+        'mahotas.cwatershed(f, markers)',
+        'pymorph.cwatershed(f, markers)',
+        'skimage.morphology.watershed(f, markers)',
[…]

### 2012-08-30 timethings.py
ENH Import image from installed directory

@@ -4,7 +4,13 @@ import skimage.feature
 import numpy as np
 import timeit
 import mahotas
-luispedro_image = '../mahotas/mahotas/demos/data/luispedro.jpg'
+
+from os import path
+luispedro_image = path.join(
+            path.dirname(mahotas.__file__),
+            'demos',
+            'data',
+            'luispedro.jpg')
 f = mahotas.imread(luispedro_image, as_grey=True)
 markers = np.zeros_like(f)
 markers[100,100] = 1

### 2012-08-30 paper.tex
TXT Add comparison with opencv

@@
 \centering
-\begin{tabular}{lrr}
+\begin{tabular}{lrrrrr}
 \toprule
-Operation & mahotas & pymorphy\\
+Operation & mahotas & pymorph & scikits-image & OpenCV\\
 \midrule
-erode       &     9.90 &    13.06 \\
-dilate      &    10.48 &     8.17 \\
-open        &    20.26 &    21.05 \\
-cwatershed  &   181.47 & 50211.56 \\
-center mass &     6.39 &       NA \\
-daubechies  &    16.92 &       NA \\
-haralick    &   276.90 &       NA \\
+erode       &     12.2 &     14.4 &     41.9 &      0.4 & \\
+dilate      &     13.4 &      9.0 &     39.4 &      0.4 & \\
+open        &     25.6 &     23.4 &     81.3 &       NA & \\
+center mass &      7.1 &       NA &       NA &       NA & \\
+sobel       &     81.9 &       NA &    106.5 &      6.2 & \\
+cwatershed  &    194.8 &  56928.5 &    288.9 &     42.9 & \\
+daubechies  &     20.7 &       NA &       NA &       NA & \\
+haralick    &    326.9 &       NA &   7497.4 &       NA & \\
 \bottomrule
 \end{tabular}
-\caption{Efficiency Results for mahotas and pymorph. Shown are values as
-multiples of the time that \code{numpy.max(f)} takes.}
+\caption{Efficiency Results for mahotas, pymorph, scikits-image, and openCV
+(through Python wrappers). Shown are values as multiples of the time that
+\code{numpy.max(f)} takes.}
 \label{tab:efficiency}
@@
 The comparison is against Pymorph~\citep{DougLotu:03}, which is a pure Python
-implementation of some of the same functions. We can see that simple
-morphological operations (erosion, dilation, opening) take a similar amount of
-time in mahotas and pymorph (pymorph intelligently uses numpy operations for
-these). However, some more complex operations such as watershed are much faster
-in mahotas.
+implementation of some of the same functions; scikits-image, which is a similar
+project to mahotas, but with a heavier emphasis on the use of
+Cython~\citep{Cython}; and OpenCV, which is a \cpp{} library with automatically
+generated Python wrappers.
+
+OpenCV is the fastest library, but this comes at the cost of some flexibility.
+Arguments to its functions must be of the exact expected type and it is
+possible to crash the interpreter if types do match the expected type (in the
+other libraries, all types are checked and an exception is generated which can
+be caught by user code). This is particularly relevant for interactive use.
+
+Among the other libraries, mahotas is the fastest, except when comparing simple
+morphological operations with pymorph. Pymorph intelligently uses arithmetic
+operations for this operation and can be very fast. However, for more complex
+methods, such as watershed; its pure Python approach is very inefficient.
 

### 2012-08-30 timethings.py

@@ -4,6 +4,7 @@ import skimage.feature
 import numpy as np
 import timeit
 import mahotas
+import cv2
 
 from os import path
 luispedro_image = path.join(
@@ -15,7 +16,7 @@ f = mahotas.imread(luispedro_image, as_grey=True)
 markers = np.zeros_like(f)
 markers[100,100] = 1
 markers[200,200] = 2
-f = f.astype(int)
+f = f.astype(np.uint8)
 markers = markers.astype(int)
 otsu = mahotas.otsu(f.astype(np.uint8))
 fbin = f > otsu
@@ -23,6 +24,10 @@ fbin8 = fbin.astype(np.uint8)
 Bc = np.eye(3)
 Bc = Bc.astype(bool)
 Bc8 = Bc.astype(np.uint8)
+f3 = np.dstack([f,f,f])
+f3 = f3.astype(np.uint8)
+f3 = f3.copy()
+markers32 = markers.astype(np.int32)
 
 pre ='''
 import skimage.filter
@@ -31,14 +36,17 @@ import skimage.feature
 import numpy as np
 import mahotas
 import pymorph
+import cv2
 import timethings
 f = timethings.f
+f3 = timethings.f3
 fbin = timethings.fbin
 fbin8 = timethings.fbin8
 f64 = f.astype(np.float64)
 Bc = timethings.Bc
 Bc8 = timethings.Bc8
 markers = timethings.markers
+markers32 = timethings.markers32
 '''
 
 def t(s):
@@ -48,20 +56,24 @@ tests = [
     ('erode', [
         'mahotas.erode(fbin, Bc)',
         'pymorph.erode(fbin, Bc)',
-        'skimage.morphology.opening(fbin8, Bc8)',
+        'skimage.morphology.erosion(fbin8, Bc8)',
+        'cv2.erode(fbin8, Bc8)',
         ]),
     ('dilate', [
         'mahotas.dilate(fbin, Bc)',
         'pymorph.dilate(fbin, Bc)',
         'skimage.morphology.dilation(fbin8, Bc8)',
+        'cv2.dilate(fbin8, Bc8)',
         ]),
     ('open', [
         'mahotas.open(fbin, Bc)',
         'pymorph.open(fbin, Bc)',
         'skimage.morphology.opening(fbin8, Bc8)',
+        None,
         ]),
     ('center mass', [
         'mahotas.center_of_mass(f)',
+        None,
         None,
         None,
         ]),
[…]

### 2012-08-31 paper.tex
TXT Credit where it is due
Add SLIC reference

@@
 
-Mahotas includes code ported and incorporated from other projects. In
-particular, the \textsc{surf} implementation is a port from the code from
+Mahotas includes code ported and incorporated from other projects. Initially,
+it was used in reproducing the functionality in the Subcellular Location Image
+Classifier (SLIC) tool from Robert F.\ Murphy's
+Lab~\citep{springerlink:10.1007/978-0-387-45524-2_47} and the initial versions
+of mahotas were designed explicitly to support that functionality. The
+\textsc{surf} implementation is a port from the code from
 \textit{dlib},\footnote{Dlib's webpage is at \url{http://dlib.net}.} a very
@@
 
+Stéfan van der Walt and Andreas Müller offered helpful comments on a draft
+version of this manuscript.
+
 \textbf{Funding}: I was supported in my work by the Funda\c c\~{a}o para a
-Ci\^encia e Tecnologia (grants SFRH/BD/37535/2007 and PTDC/SAU-GMG/115652/2008)
-and by a grant from the Siebel Scholars Foundation.
+Ci\^encia e Tecnologia (grants SFRH/BD/37535/2007 and
+PTDC/SAU-GMG/115652/2008), by NIH grant GM078622 (to Robert F.\ Murphy), by a
+grant from the Scaife Foundation, by the HHMI Interface initial, and by a grant
+from the Siebel Scholars Foundation.
 

### 2012-08-31 timethings.py
MIN Slightly better timing

@@ -50,7 +50,7 @@ markers32 = timethings.markers32
 '''
 
 def t(s):
-    return timeit.timeit(s, setup=pre, number=24)
+    return min(timeit.timeit(s, setup=pre, number=24) for i in xrange(3))
 
 tests = [
     ('erode', [

### 2012-08-31 paper.tex
TXT Update timings after mahotas update

@@
 \midrule
-erode       &     12.2 &     14.4 &     41.9 &      0.4 & \\
-dilate      &     13.4 &      9.0 &     39.4 &      0.4 & \\
-open        &     25.6 &     23.4 &     81.3 &       NA & \\
+erode       &     10.3 &     14.2 &     41.5 &      0.4 & \\
+dilate      &     12.7 &      8.9 &     38.9 &      0.4 & \\
+open        &     23.0 &     23.3 &     80.8 &       NA & \\
 center mass &      7.1 &       NA &       NA &       NA & \\
-sobel       &     81.9 &       NA &    106.5 &      6.2 & \\
-cwatershed  &    194.8 &  56928.5 &    288.9 &     42.9 & \\
-daubechies  &     20.7 &       NA &       NA &       NA & \\
-haralick    &    326.9 &       NA &   7497.4 &       NA & \\
+sobel       &     75.4 &       NA &    105.4 &      6.2 & \\
+cwatershed  &    191.3 &  60410.0 &    285.0 &     42.1 & \\
+daubechies  &     20.3 &       NA &       NA &       NA & \\
+haralick    &    304.5 &       NA &   7615.8 &       NA & \\
 \bottomrule
@@
 (through Python wrappers). Shown are values as multiples of the time that
-\code{numpy.max(f)} takes.}
+\code{numpy.max(image)} takes to compute the maximum pixel value in the image
+(all operations are over the same image). For scikits-image, features on the
+grey-scale cooccurrence matrix were used instead of Haralick features, which it
+does not support.}
 \label{tab:efficiency}
@@
 other libraries, all types are checked and an exception is generated which can
-be caught by user code). This is particularly relevant for interactive use.
+be caught by user code). This is particularly relevant for interactive use as
+the user is often exploring and is willing to pay the cost of a few extra type
+conversions to avoid a hard-crash.
 

### 2012-09-01 paper.tex
TXT Readthrough fixes

@@
 Mahotas is a computer vision library for the Python Programming Language
-(versions 2.5 and up, including version 3). It operates on numpy
+(versions 2.5 and up, including version 3 and up). It operates on numpy
 arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
@@
 Java~\citep{Marcel:2010:TMP:1873951.1874254}, Mahotas does not need to define a
-data structure, but uses the numpy array structure. Many basic manipulation
-functionality that would otherwise be part of a computer vision library are
-handled by numpy, for example computing averages and other simple statistics,
-handling multi-channel images, converting between types (integer and floating
-point images are supported by mahotas, whenever it is meaningful). For the
-user, this has the additional advantage that they do not need to learn yet
-another set of functions.
+new image data structure, but uses the numpy array structure. Many basic
+manipulation functionality that would otherwise be part of a computer vision
+library are handled by numpy, for example computing averages and other simple
+statistics, handling multi-channel images, converting between types (integer
+and floating point images are supported by mahotas). For the user, this has the
+additional advantage that they do not need to learn yet another set of
+functions.
 
@@
 well-tested package (the first version was made available over 4~years ago).
-\footnote{Note for reviewers: the version currently available implements all
-functionality described in this manuscript. I will release it as version 1.0
-when this manuscript is accepted to coincide with publication. Naturally, any
-bugs that are found and reported in the meanwhile will be addressed.} Mahotas
-runs and is used on different versions of Unix (including Linux, SunOS, and
-FreeBSD), Mac OS X, and Windows.
+\footnote{Note for reviewers: the version currently available (0.9.2)
+implements all functionality described in this manuscript. I will release it as
+version 1.0 when this manuscript is accepted to coincide with publication.
+Naturally, any bugs that are found and reported in the meanwhile will be
+addressed.} Mahotas runs and is used on different versions of Unix (including
+Linux, SunOS, and FreeBSD), Mac OS X, and Windows.
 
@@
 transforms~\citep{felzenszwalb}.
-\item[filtering] Gaussian filtering and general convolutions.
+\item[filtering] Gaussian filtering, edge finding, and general convolutions.
 \item[polygon operations] convex hull, polygon drawing.
@@
 where the output will be stored. This argument is often much more restricted in
-type. In particular, it must often be a contiguous array.\footnote{Numpy
-supports non-contiguous arrays, which are most often slices into other, larger,
+type. In particular, it must be a contiguous array.\footnote{Numpy supports
+non-contiguous arrays, which are most often slices into other, larger,
 contiguous arrays (e.g., given a $128 \times 128$ contiguous array, one can
@@
         !numpy::are_arrays(array, Bc, output) || !numpy::same_shape(array, output) ||
-        !PyArray_EquivTypenums(PyArray_TYPE(array), PyArray_TYPE(Bc)) ||
-        !PyArray_EquivTypenums(PyArray_TYPE(array), PyArray_TYPE(output)) ||
+        !numpy::equiv_typenums(array, Bc, output) ||
         PyArray_NDIM(array) != PyArray_NDIM(Bc)
@@
 specialisation of the template that does the actual work. In this example
-\code{erode} implements (binary) erosion:
+\code{erode} implements erosion:
 
@@
 is correctly released even if the scope is left through an exception
-\citep{Stroustrup1994}.} object that release the Python global interpreter lock
+\citep{Stroustrup1994}.} object that releases the Python global interpreter lock
 (\textsc{gil})\footnote{In the CPython interpreter, the most commonly used
@@
 mahotas functions to run concurrently.
-\item[\code{array}] This is a thin wrapper around \code{PyArrayObject} that
-knows its data type and has iterators which resemble the \cpp{} standard
-library. This makes the code type-safer. This is also a \textsc{raii} object in
-terms of managing Python reference counts.
+\item[\code{array}] This is a thin wrapper around \code{PyArrayObject}, the raw
+numpy data type, which has iterators which resemble the \cpp{} standard
+library. It also handles type-casting internally, making the code type-safer.
+This is also a \textsc{raii} object in terms of managing Python reference
+counts.
 \item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it
@@
 normalized to multiples of the time it takes to go over the image and find its
-maximum pixel value (this was done using \code{numpy.max(image}). The
+maximum pixel value (this was done using \code{numpy.max(image)}). The
 measurements shown were obtained on an Intel 64~bit system, running Ubuntu
@@
 be caught by user code). This is particularly relevant for interactive use as
-the user is often exploring and is willing to pay the cost of a few extra type
-conversions to avoid a hard-crash.
+the user is often exploring and is willing to pay the speed cost of a few extra
+type conversions to avoid a hard-crash.
 
@@
 \textbf{Additional system requirements}\\
-None.
+None at runtime. Compilation from source requires a \cpp{} compiler.
 
@@
 \textit{Licence}: MIT\\
-\textit{Date published}: Since 2009
+\textit{Date published}: Since 2010
 
@@
 PTDC/SAU-GMG/115652/2008), by NIH grant GM078622 (to Robert F.\ Murphy), by a
-grant from the Scaife Foundation, by the HHMI Interface initial, and by a grant
-from the Siebel Scholars Foundation.
+grant from the Scaife Foundation, by the HHMI Interfaces Initiative, and by a
+grant from the Siebel Scholars Foundation.
 

### 2012-09-01 paper.tex
MIN Make sure 'van der Walt' is correct

@@
     natbib=true,%
+    useprefix=true,
     backend=biber,%

### 2012-09-04 paper.tex
MIN Fit affiliations in one line
TXT The missing patch-submitter

@@
 \author{Luis Pedro Coelho\\
-Lane Center for Computational Biology, Carnegie Mellon University, Pittsburgh, USA\\
-Instituto de Medicina Molecular, Lisboa, Portugal}
+Lane Center for Computational Biology, Carnegie Mellon University\\
+Instituto de Medicina Molecular}
 
@@
 Joe Kington (University of Wisconsin), Christoph Gohlke (University of
-California, Irvine), and Sandro Knauss (University of Bremen).
+California, Irvine), Lukas Bossard (ETH Zurich), and Sandro Knauss (University
+of Bremen).
 

### 2012-11-20 paper.tex
MIN Breakup lines to fit better on page
STYL Adjust look for arXiv

@@
-\documentclass{article}
+\documentclass{scrartcl}
 \usepackage{hyperref}
@@
 \usepackage{python}
+\usepackage{fontspec}
 \usepackage[%
@@
     colorlinks=true,
+    pdftitle={Mahotas: Open source software for scriptable computer vision},
     pdfauthor={Luis Pedro Coelho},
@@
 }
+\font\CodeFont="DejaVu Sans Mono" at 8pt
+\setmainfont[Mapping=tex-text]{Linux Libertine O}
 \newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
-\let\code\texttt
+\def\code#1{\bgroup\CodeFont #1\egroup}
 \addbibresource{references.bib}
@@
 Instituto de Medicina Molecular}
+\date{September 2012}
 
@@
 
-Mahotas is available under a liberal open source license (MIT License) and is
-available from \url{http://github.com/luispedro/mahotas} or the Python Package
+Mahotas is released under a liberal open source license (MIT License) and is
+available from \url{http://github.com/luispedro/mahotas} and from the Python Package
 Index (\url{http://pypi.python.org/pypi/mahotas}).
@@
     if (!PyArg_ParseTuple(args, "OOO", &array, &Bc, &output) ||
-        !numpy::are_arrays(array, Bc, output) || !numpy::same_shape(array, output) ||
+        !numpy::are_arrays(array, Bc, output) ||
+        !numpy::same_shape(array, output) ||
         !numpy::equiv_typenums(array, Bc, output) ||
@@
     typename numpy::aligned_array<T>::iterator iter = array.begin();
-    filter_iterator<T> filter(array.raw_array(), Bc.raw_array(), EXTEND_NEAREST, is_bool(T()));
+    filter_iterator<T> filter(array.raw_array(), Bc.raw_array(),
+                    EXTEND_NEAREST, is_bool(T()));
     const int N2 = filter.size();

### 2012-11-21 paper.tex
BLD Use pdflatex (which arXiv uses)

@@
 \usepackage{python}
-\usepackage{fontspec}
-\usepackage[%
-    natbib=true,%
-    useprefix=true,
-    backend=biber,%
-    backref=true,%
-    citestyle=authoryear-comp,%
-    bibstyle=authoryear,%
-    maxbibnames=24,%
-    maxcitenames=2,%
-]{biblatex}
-
+\usepackage{libertine}
+\usepackage{DejaVuSansMono}
+\usepackage{natbib}
 \hypersetup{%
@@
 }
-\font\CodeFont="DejaVu Sans Mono" at 8pt
-\setmainfont[Mapping=tex-text]{Linux Libertine O}
+
 \newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
-\def\code#1{\bgroup\CodeFont #1\egroup}
-\addbibresource{references.bib}
+\let\code\texttt
+\bibliographystyle{natbib}
 
@@
 bit of initialisation, and finally, switches in the input type with the help of
-the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES()} macro, which call the right
+the \code{SAFE\_\-SWITCH\_\-ON\_\-INTEGER\_\-TYPES()} macro, which call the right
 specialisation of the template that does the actual work. In this example
@@
 However, all of the Python interface code is hand-written, as opposed to using
-automatic interface generators like Swig~\cite{Beazley2003599}. This is more
+automatic interface generators like Swig~\citep{Beazley2003599}. This is more
 work, but the end result is of much higher quality, especially when it comes to
@@
 
-\printbibliography
+\bibliography{references}
 \end{document}

### 2012-11-21 build.sh

@@ -4,8 +4,8 @@ function build() {
   input=$1
   mkdir -p .$input.tex_files
   cd .$input.tex_files
-  TEXINPUTS=.:..:../images/:../figures/: xelatex $input
-  BSTINPUTS=:..: BIBINPUTS=:..:.: biber $input
+  TEXINPUTS=.:..:../images/:../figures/: pdflatex $input
+  BSTINPUTS=:..: BIBINPUTS=:..:.: bibtex $input
   cp $input.pdf ..
   cd ..
 }

### 2012-12-06 paper.tex
TXT Improve text. Update download numbers

--- 2012-11-20 paper.tex
@@
 
-Additionally, by integrating into the Python numeric ecosystem, users can use
-other packages in a seamless way. In particular, mahotas does not implement any
-machine learning functionality, but rather advises the user to use another,
-specialised package, such as scikits-learn or milk.
-
 Python is a natural ``glue'' language: it is easy to use state-of-the-art
@@
 
+Additionally, by integrating into the Python numeric ecosystem, users can use
+other packages in a seamless way. In particular, mahotas does not implement any
+machine learning functionality, but rather advises the user to use another,
+specialised, package, such as scikits-learn or milk.
+
 Version 1.0 of mahotas has been released recently and this is now a mature,
 well-tested package (the first version was made available over 4~years ago).
-\footnote{Note for reviewers: the version currently available (0.9.2)
+\footnote{Note for reviewers: the version currently available (0.9.6)
 implements all functionality described in this manuscript. I will release it as
@@
 
-Mahotas is mostly written in \cpp, but this is completely hidden from the user
-as there are hand-written Python wrappers for all functions (automatically
-generated wrappers inevitably lead to worse error messages and are less
-flexible).
+The majority of functions in mahotas have an implementation in \cpp, and a
+hand-written Python wrapper. Automatically generated wrappers inevitably lead
+to worse error messages and are less flexible.
 
@@
                 numpy::aligned_array<type>(Bc));
-    SAFE_SWITCH_ON_INTEGER_TYPES_OF(array, true);
+    SAFE_SWITCH_ON_INTEGER_TYPES_OF(array);
 #undef HANDLE
@@
 the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES()} macro, which call the right
-specialisation of the template that does the actual work. In this example
-\code{erode} implements erosion:
+specialisation of the template that does the actual work. These macros will
+also take care of catching any \cpp{} exceptions and converting them to Python
+exceptions. In this example, \code{erode} implements erosion:
 
@@
 
-Among the other libraries, mahotas is the fastest, except when comparing simple
-morphological operations with pymorph. Pymorph intelligently uses arithmetic
-operations for this operation and can be very fast. However, for more complex
-methods, such as watershed; its pure Python approach is very inefficient.
+Among the other libraries, mahotas is the fastest, except for the case of a
+large median filter. For simple operations, pymorph intelligently uses
+arithmetic operations for this operation and can be very fast. However, for
+more complex methods, such as watershed; its pure Python approach is very
+inefficient. Skimage also uses compiled code to achieve fast results.
 
@@
 
-Mahotas includes a complete automated suite of unit tests. These all
+Mahotas includes a complete automated suite of unit tests. These test all
 functionality and include several regression tests. There are no known bugs in
@@
 Mahotas has been available in the Python Package Index since April~2010 and has
-been downloaded over 20,000~times. This does not include any downloads from
+been downloaded over 35,000~times. This does not include any downloads from
 other sources. Mahotas includes a full test suite. There are no known bugs.

### 2012-12-29 paper.tex
TXT Text fixes and updates

--- 2012-11-21 paper.tex
@@
-\documentclass{scrartcl}
+\documentclass{article}
 \usepackage{hyperref}
@@
 \usepackage{python}
-\usepackage{libertine}
-\usepackage{DejaVuSansMono}
-\usepackage{natbib}
+\usepackage[%
+    natbib=true,%
+    useprefix=true,
+    backend=biber,%
+    backref=true,%
+    citestyle=authoryear-comp,%
+    bibstyle=authoryear,%
+    maxbibnames=24,%
+    maxcitenames=2,%
+]{biblatex}
+
 \hypersetup{%
@@
     colorlinks=true,
-    pdftitle={Mahotas: Open source software for scriptable computer vision},
     pdfauthor={Luis Pedro Coelho},
@@
 }
-
 \newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
 \let\code\texttt
-\bibliographystyle{natbib}
+\addbibresource{references.bib}
 
@@
 Instituto de Medicina Molecular}
-\date{September 2012}
 
@@
 
-Mahotas is released under a liberal open source license (MIT License) and is
-available from \url{http://github.com/luispedro/mahotas} and from the Python Package
+Mahotas is available under a liberal open source license (MIT License) and is
+available from \url{http://github.com/luispedro/mahotas} or the Python Package
 Index (\url{http://pypi.python.org/pypi/mahotas}).
@@
 manipulation functionality that would otherwise be part of a computer vision
-library are handled by numpy, for example computing averages and other simple
+library are handled by numpy. For example, computing averages and other simple
 statistics, handling multi-channel images, converting between types (integer
-and floating point images are supported by mahotas). For the user, this has the
-additional advantage that they do not need to learn yet another set of
-functions.
+and floating point images are supported by mahotas) can all be performed with
+numpy builtin functionality. For the user, this has the additional advantage
+that they do not need to learn yet another set of functions.
 
@@
 well-tested package (the first version was made available over 4~years ago).
-\footnote{Note for reviewers: the version currently available (0.9.2)
+\footnote{Note for reviewers: the version currently available (0.9.6)
 implements all functionality described in this manuscript. I will release it as
@@
 versions~\citep{Coelho2010}).
-\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
+\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms are
+supported.
 \item[morphological functions] Erosion and dilation, as well as some more
@@
 Generally, the default is to use a $3 \times 3$ cross as the default if no
-structuring filter is given (the exception to this rule is the median filter,
-where the default is a $3 \times 3$ square).
+structuring filter is given.
 
@@
 
-This is a simple example of mahotas usage. Code for this and other examples is
-present in the mahotas source distribution under the \texttt{demos/} directory.
-In this example, we load an image and find SURF interest point and descriptors
-on it.
+Code for this and other examples is present in the mahotas source distribution
+under the \texttt{demos/} directory. In this example, we load an image and find
+SURF interest point and descriptors on it.
 
@@
 
-Finally, we can show the points in colour.
+Finally, we can show the points in different colours.
 \begin{python}
@@
 The easy interaction with matplotlib is another way in which we benefit from
-the numpy-based ecosystem.
+the numpy-based ecosystem as mahotas does not need to support interacting with
+a graphical system to display images.
 
@@
 bit of initialisation, and finally, switches in the input type with the help of
-the \code{SAFE\_\-SWITCH\_\-ON\_\-INTEGER\_\-TYPES()} macro, which call the right
+the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES} macro, which call the right
 specialisation of the template that does the actual work. In this example
@@
     filter_iterator<T> filter(array.raw_array(), Bc.raw_array(),
-                    EXTEND_NEAREST, is_bool(T()));
+                    ExtendNearest, is_bool(T()));
     const int N2 = filter.size();
@@
 compiler instantiates the function for the several integer and floating point
-types. Given the small size of these functions, this is not a big issue.
+types. Given the small size of these functions, the total size of the compiled
+library is reasonable (circa 6MiB on an Intel-based 64~bit system).
 
@@
 This is also a \textsc{raii} object in terms of managing Python reference
-counts.
-\item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it
-is useful to iterate over an image and use a centered filter around each pixel
-(it keeps track of all of the boundary conditions).
+counts. In mahotas debug builds, this object additionally adds several checks
+to all the memory acesses.
+\item[\code{filter\_iterator}] This was adapted from code in the
+\code{scipy.ndimage} packages and it is useful to iterate over an image and use
+a centered filter around each pixel (it keeps track of all of the boundary
+conditions).
 \end{description}
@@
 possible to crash the interpreter if types do match the expected type (in the
-other libraries, all types are checked and an exception is generated which can
-be caught by user code). This is particularly relevant for interactive use as
-the user is often exploring and is willing to pay the speed cost of a few extra
-type conversions to avoid a hard-crash.
+other libraries, including mahotas, all types are checked and an exception is
+generated which can be caught by user code). This is particularly relevant for
+interactive use as the user is often exploring and is willing to pay the speed
+cost of a few extra type conversions to avoid a hard-crash.
 
-Among the other libraries, mahotas is the fastest, except when comparing simple
-morphological operations with pymorph. Pymorph intelligently uses arithmetic
-operations for this operation and can be very fast. However, for more complex
-methods, such as watershed; its pure Python approach is very inefficient.
+Among the other libraries, mahotas is the fastest. Pymorph, even though it is
+implemented in Python only, intelligently uses arithmetic operations for this
+operation and can be very fast. However, for more complex methods, such as
+watershed; its pure Python approach is very inefficient.
 
@@
 
-Mahotas includes a complete automated suite of unit tests. These all
+Mahotas includes a complete automated suite of unit tests, which tests all
 functionality and include several regression tests. There are no known bugs in
@@
 Mahotas does not include machine learning related functionality, such as
-$k$-means clustering or any classification. This is the result of an explicit
-design decision. Specialised machine learning packages for Python already
+$k$-means clustering or classification methods. This is the result of an
+explicit design decision. Specialised machine learning packages for Python
+already
 exist~\citep{Pedregosa:2011:SML:2078183.2078195,springerlink:10.1007/978-3-540-30116-5_58,Schaul:2010:PYB:1756006.1756030,Sonnenburg:2010:SML:1756006.1859911}.
@@
 automatic interface generators like Swig~\citep{Beazley2003599}. This is more
-work, but the end result is of much higher quality, especially when it comes to
-giving useful error messages (e.g., when a type mismatch occurs, an automatic
-system will often be forced to resort to a generic message as it does not have
-any knowledge of what the arguments mean besides their automatically inferred
-types).
+work intially, but the end result is of much higher quality, especially when it
+comes to giving useful error messages (e.g., when a type mismatch occurs, an
+automatic system will often be forced to resort to a generic message as it does
+not have any knowledge of what the arguments mean besides their automatically
+inferred types).
 
 Mahotas has been available in the Python Package Index since April~2010 and has
-been downloaded over 20,000~times. This does not include any downloads from
+been downloaded over 40,000~times. This does not include any downloads from
 other sources. Mahotas includes a full test suite. There are no known bugs.
@@
 
-\bibliography{references}
+\printbibliography
 \end{document}

### 2012-12-29 paper.tex
BUG Clarify dates of release

@@
 Version 1.0 of mahotas has been released recently and this is now a mature,
-well-tested package (the first version was made available over 4~years ago).
-\footnote{Note for reviewers: the version currently available (0.9.6)
-implements all functionality described in this manuscript. I will release it as
-version 1.0 when this manuscript is accepted to coincide with publication.
-Naturally, any bugs that are found and reported in the meanwhile will be
-addressed.} Mahotas runs and is used on different versions of Unix (including
-Linux, SunOS, and FreeBSD), Mac OS X, and Windows.
+well-tested package (the first versions were made available over 4~years ago,
+although the package was not named mahotas then). \footnote{Note for reviewers:
+the version currently available (0.9.6) implements all functionality described
+in this manuscript. I will release it as version 1.0 when this manuscript is
+accepted to coincide with publication. Naturally, any bugs that are found and
+reported in the meanwhile will be addressed.} Mahotas runs and is used on
+different versions of Unix (including Linux, SunOS, and FreeBSD), Mac OS X, and
+Windows.
 
@@
 \textit{Licence}: MIT\\
-\textit{Date published}: Since 2010
+\textit{Date published}: Since 2010. Some of the code had been previously made
+available under other names.
 

### 2013-01-02 paper.tex
TXT Update performance table

@@
 
-So, for example, this is how \emph{erode} is implemented. \code{py\_erode}
-consists mostly of boiler-plate code:
+So, for example, this is how \emph{erode} is implemented.\footnote{This is the
+generic version of the code. Erode, like a few other functions, has two
+versions, a fast version, limited to two-dimensional, contiguous, images and a
+generic one presented here.} \code{py\_erode} consists mostly of boiler-plate
+code:
 
@@
 \midrule
-erode       &     10.3 &     14.2 &     41.5 &      0.4 & \\
-dilate      &     12.7 &      8.9 &     38.9 &      0.4 & \\
-open        &     23.0 &     23.3 &     80.8 &       NA & \\
-center mass &      7.1 &       NA &       NA &       NA & \\
-sobel       &     75.4 &       NA &    105.4 &      6.2 & \\
-cwatershed  &    191.3 &  60410.0 &    285.0 &     42.1 & \\
-daubechies  &     20.3 &       NA &       NA &       NA & \\
-haralick    &    304.5 &       NA &   7615.8 &       NA & \\
+erode                &      1.6 &     15.1 &      7.4 &      0.4 & \\
+dilate               &      1.5 &      9.1 &      7.3 &      0.4 & \\
+open                 &      3.2 &     24.3 &     14.8 &       NA & \\
+median filter (2)    &    226.9 &       NA &   2034.0 &       NA & \\
+median filter (10)   &   2810.9 &       NA &   1877.1 &       NA & \\
+center mass          &      5.0 &       NA &   3611.2 &       NA & \\
+sobel                &     34.1 &       NA &     62.5 &      6.2 & \\
+cwatershed           &    174.8 &  58440.3 &    287.3 &     44.9 & \\
+daubechies           &     18.8 &       NA &       NA &       NA & \\
+haralick             &    233.1 &       NA &   7760.7 &       NA & \\
 \bottomrule
@@
 grey-scale cooccurrence matrix were used instead of Haralick features, which it
-does not support.}
+does not support. In the case of \emph{median filter}, the radius of the
+structuring element is shown in parentheses.}
 \label{tab:efficiency}
@@
 operation and can be very fast. However, for more complex methods, such as
-watershed; its pure Python approach is very inefficient.
+watershed; its pure Python approach is very inefficient. The one exception is
+that median filtering with a large structuring element is faster in
+scikit-image. In fact, that library uses an algorithm with better asymptotic
+behavior for this operation.
 
@@
 \textbf{Programming language}\\
-Mahotas works in Python (minimal version is~2.5 and works in all more recent
-versions, including in the Python~3 series).
+Mahotas works in Python (minimal version is~2.5, but mahotas works with all
+more recent versions, including version in the Python~3 series).
 
@@
 \textit{Licence}: MIT\\
-\textit{Date published}: Since 2010. Some of the code had been previously made
-available under other names.
+\textit{Date published}: Since 2010 as mahotas. Some of the code had been
+previously made available under other names.
 

### 2013-01-02 paper.tex

--- 2012-12-29 paper.tex
@@
 \usepackage{python}
-\usepackage[%
-    natbib=true,%
-    useprefix=true,
-    backend=biber,%
-    backref=true,%
-    citestyle=authoryear-comp,%
-    bibstyle=authoryear,%
-    maxbibnames=24,%
-    maxcitenames=2,%
-]{biblatex}
-
+\usepackage{libertine}
+\usepackage{DejaVuSansMono}
+\usepackage{natbib}
 \hypersetup{%
@@
 }
+
 \newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
 \let\code\texttt
-\addbibresource{references.bib}
-
+\bibliographystyle{natbib}
 \title{Mahotas: Open source software for scriptable computer vision}
@@
 manipulation functionality that would otherwise be part of a computer vision
-library are handled by numpy. For example, computing averages and other simple
+library are handled by numpy, for example computing averages and other simple
 statistics, handling multi-channel images, converting between types (integer
-and floating point images are supported by mahotas) can all be performed with
-numpy builtin functionality. For the user, this has the additional advantage
-that they do not need to learn yet another set of functions.
-
-Additionally, by integrating into the Python numeric ecosystem, users can use
-other packages in a seamless way. In particular, mahotas does not implement any
-machine learning functionality, but rather advises the user to use another,
-specialised package, such as scikits-learn or milk.
+and floating point images are supported by mahotas). For the user, this has the
+additional advantage that they do not need to learn yet another set of
+functions.
 
@@
 balance between speed and ease of implementation.
+
+Additionally, by integrating into the Python numeric ecosystem, users can use
+other packages in a seamless way. In particular, mahotas does not implement any
+machine learning functionality, but rather advises the user to use another,
+specialised, package, such as scikits-learn or milk.
 
@@
 versions~\citep{Coelho2010}).
-\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms are
-supported.
+\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
 \item[morphological functions] Erosion and dilation, as well as some more
@@
 Generally, the default is to use a $3 \times 3$ cross as the default if no
-structuring filter is given.
+structuring filter is given (the exception to this rule is the median filter,
+where the default is a $3 \times 3$ square).
 
@@
 
-Code for this and other examples is present in the mahotas source distribution
-under the \texttt{demos/} directory. In this example, we load an image and find
-SURF interest point and descriptors on it.
+This is a simple example of mahotas usage. Code for this and other examples is
+present in the mahotas source distribution under the \texttt{demos/} directory.
+In this example, we load an image and find SURF interest point and descriptors
+on it.
 
@@
 
-Finally, we can show the points in different colours.
+Finally, we can show the points in colour.
 \begin{python}
@@
 The easy interaction with matplotlib is another way in which we benefit from
-the numpy-based ecosystem as mahotas does not need to support interacting with
-a graphical system to display images.
+the numpy-based ecosystem.
 
@@
 bit of initialisation, and finally, switches in the input type with the help of
-the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES} macro, which call the right
+the \code{SAFE\_\-SWITCH\_\-ON\_\-INTEGER\_\-TYPES()} macro, which call the right
 specialisation of the template that does the actual work. In this example
@@
     filter_iterator<T> filter(array.raw_array(), Bc.raw_array(),
-                    ExtendNearest, is_bool(T()));
+                    EXTEND_NEAREST, is_bool(T()));
     const int N2 = filter.size();
@@
 compiler instantiates the function for the several integer and floating point
-types. Given the small size of these functions, the total size of the compiled
-library is reasonable (circa 6MiB on an Intel-based 64~bit system).
+types. Given the small size of these functions, this is not a big issue.
 
@@
 This is also a \textsc{raii} object in terms of managing Python reference
-counts. In mahotas debug builds, this object additionally adds several checks
-to all the memory acesses.
-\item[\code{filter\_iterator}] This was adapted from code in the
-\code{scipy.ndimage} packages and it is useful to iterate over an image and use
-a centered filter around each pixel (it keeps track of all of the boundary
-conditions).
+counts.
+\item[\code{filter\_iterator}] This is taken from \code{scipy.ndimage} and it
+is useful to iterate over an image and use a centered filter around each pixel
+(it keeps track of all of the boundary conditions).
 \end{description}
@@
 possible to crash the interpreter if types do match the expected type (in the
-other libraries, including mahotas, all types are checked and an exception is
-generated which can be caught by user code). This is particularly relevant for
-interactive use as the user is often exploring and is willing to pay the speed
-cost of a few extra type conversions to avoid a hard-crash.
+other libraries, all types are checked and an exception is generated which can
+be caught by user code). This is particularly relevant for interactive use as
+the user is often exploring and is willing to pay the speed cost of a few extra
+type conversions to avoid a hard-crash.
 
-Among the other libraries, mahotas is the fastest. Pymorph, even though it is
-implemented in Python only, intelligently uses arithmetic operations for this
-operation and can be very fast. However, for more complex methods, such as
-watershed; its pure Python approach is very inefficient.
+Among the other libraries, mahotas is the fastest, except for the case of a
+large median filter. For simple operations, pymorph intelligently uses
+arithmetic operations for this operation and can be very fast. However, for
+more complex methods, such as watershed; its pure Python approach is very
+inefficient. Skimage also uses compiled code to achieve fast results.
 
@@
 
-Mahotas includes a complete automated suite of unit tests, which tests all
+Mahotas includes a complete automated suite of unit tests. These test all
 functionality and include several regression tests. There are no known bugs in
@@
 Mahotas does not include machine learning related functionality, such as
-$k$-means clustering or classification methods. This is the result of an
-explicit design decision. Specialised machine learning packages for Python
-already
+$k$-means clustering or any classification. This is the result of an explicit
+design decision. Specialised machine learning packages for Python already
 exist~\citep{Pedregosa:2011:SML:2078183.2078195,springerlink:10.1007/978-3-540-30116-5_58,Schaul:2010:PYB:1756006.1756030,Sonnenburg:2010:SML:1756006.1859911}.
@@
 automatic interface generators like Swig~\citep{Beazley2003599}. This is more
-work intially, but the end result is of much higher quality, especially when it
-comes to giving useful error messages (e.g., when a type mismatch occurs, an
-automatic system will often be forced to resort to a generic message as it does
-not have any knowledge of what the arguments mean besides their automatically
-inferred types).
+work, but the end result is of much higher quality, especially when it comes to
+giving useful error messages (e.g., when a type mismatch occurs, an automatic
+system will often be forced to resort to a generic message as it does not have
+any knowledge of what the arguments mean besides their automatically inferred
+types).
 
 Mahotas has been available in the Python Package Index since April~2010 and has
-been downloaded over 40,000~times. This does not include any downloads from
+been downloaded over 35,000~times. This does not include any downloads from
 other sources. Mahotas includes a full test suite. There are no known bugs.

### 2013-01-03 paper.tex

--- 2013-01-02 paper.tex
@@
 }
+
 \newcommand*{\cpp}{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\textbf{++}}}}
 \let\code\texttt
-\addbibresource{references.bib}
+\bibliographystyle{natbib}
 

### 2013-01-05 paper.tex
MIN Update date
TXT Corrections from a read through

@@
 Instituto de Medicina Molecular}
-\date{September 2012}
+\date{January 2013}
 
@@
 appropriate for fast development, but the algorithms are implemented in \cpp{}
-and are tuned for speed.
+and are tuned for speed. The library is designed to fit in with the scientific
+software ecosystem in this language and can leverage the existing
+infrastructure developed in that language.
 
@@
 
-Additionally, by integrating into the Python numeric ecosystem, users can use
-other packages in a seamless way. In particular, mahotas does not implement any
-machine learning functionality, but rather advises the user to use another,
-specialised package, such as scikits-learn or milk.
+It contains over 100~functions with functionality ranging from traditional
+image filtering and morphological operations to more modern wavelet
+decompositions and local feature computations. Additionally, by integrating
+into the Python numeric ecosystem, users can use other packages in a seamless
+way. In particular, mahotas does not implement any machine learning
+functionality, but rather advises the user to use another, specialised package,
+such as scikits-learn or milk.
 
@@
 \begin{description}
-\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This
+\item[\textsc{Surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This
 includes both keypoint detection and descriptor computation.
-\item[features] Global feature descriptors. In particular, Haralick texture
+\item[Features] Global feature descriptors. In particular, Haralick texture
 features, Zernike moments, local binary patterns, and threshold adjacency
@@
 versions~\citep{Coelho2010}).
-\item[wavelet] Haar and Daubechies wavelets. Forward and inverse transforms are
+\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms are
 supported.
-\item[morphological functions] Erosion and dilation, as well as some more
+\item[Morphological functions] Erosion and dilation, as well as some more
 complex operations built on these. There are both binary and grayscale
 implementations of these operators.
-\item[watershed] seeded watershed and distance map
+\item[Watershed] seeded watershed and distance map
 transforms~\citep{felzenszwalb}.
-\item[filtering] Gaussian filtering, edge finding, and general convolutions.
-\item[polygon operations] convex hull, polygon drawing.
+\item[Filtering] Gaussian filtering, edge finding, and general convolutions.
+\item[Polygon operations] convex hull, polygon drawing.
 \end{description}
 
-In general, functions operate on all data types. This is performed without any
+Numpy arrays contain data of a specific type, such \code{unsigned 8 bit
+integer} or floating point numbers. While natural colour images are typically
+8~bits, scientific data is often larger and processing can result in floating
+point images. Mahotas works on all datatypes. This is performed without any
 extra memory copies. Mahotas is heavily optimised for both speed and memory
@@
 Code for this and other examples is present in the mahotas source distribution
-under the \texttt{demos/} directory. In this example, we load an image and find
-SURF interest point and descriptors on it.
+under the \texttt{demos/} directory. In this example, we load an image, find
+SURF interest points, and compute descriptors.
 
@@
 
-Using numpy operations and milk, we can select only the descriptors and cluster
-them into a group of five:
+The \code{surf.surf} function returns both the descriptors and their meta data.
+We use numpy operations to retain only the descriptors (the meta data is in the
+first five positions):
 
@@
 descrs = spoints[:,6:]
+\end{python}
+
+Using milk, we cluster the descriptors into five groups:
+
+\begin{python}
 values, _ = milk.kmeans(descrs, 5)
@@
 Mahotas is mostly written in \cpp, but this is completely hidden from the user
-as there are hand-written Python wrappers for all functions (automatically
+as there are hand-written Python wrappers for all functions. Automatically
 generated wrappers inevitably lead to worse error messages and are less
-flexible).
+flexible.
 
@@
 versions, a fast version, limited to two-dimensional, contiguous, images and a
-generic one presented here.} \code{py\_erode} consists mostly of boiler-plate
-code:
+generic one presented here. The selection between the two implementations is
+done automatically for the user.} \code{py\_erode} consists mostly of
+boiler-plate code:
 
@@
 bit of initialisation, and finally, switches in the input type with the help of
-the \code{SAFE\_SWITCH\_ON\_INTEGER\_TYPES} macro, which call the right
+the \code{SAFE\_\-SWITCH\_\-ON\_\-INTEGER\_\-TYPES} macro, which call the right
 specialisation of the template that does the actual work. In this example
@@
 types. Given the small size of these functions, the total size of the compiled
-library is reasonable (circa 6MiB on an Intel-based 64~bit system).
+library is reasonable (circa 6MiB on an Intel-based 64~bit system for the whole
+library).
 
@@
 \centering
-\begin{tabular}{lrrrrr}
+\begin{tabular}{lrrrrrr}
 \toprule
-Operation & mahotas & pymorph & scikits-image & OpenCV\\
+Operation            &  mahotas &  pymorph & scikits-image & OpenCV \\
 \midrule
-erode                &      1.6 &     15.1 &      7.4 &      0.4 & \\
-dilate               &      1.5 &      9.1 &      7.3 &      0.4 & \\
-open                 &      3.2 &     24.3 &     14.8 &       NA & \\
-median filter (2)    &    226.9 &       NA &   2034.0 &       NA & \\
-median filter (10)   &   2810.9 &       NA &   1877.1 &       NA & \\
-center mass          &      5.0 &       NA &   3611.2 &       NA & \\
-sobel                &     34.1 &       NA &     62.5 &      6.2 & \\
-cwatershed           &    174.8 &  58440.3 &    287.3 &     44.9 & \\
-daubechies           &     18.8 &       NA &       NA &       NA & \\
-haralick             &    233.1 &       NA &   7760.7 &       NA & \\
+erode                &      1.6 &     15.1 &      7.4 &      0.4  \\
+dilate               &      1.5 &      9.1 &      7.3 &      0.4  \\
+open                 &      3.2 &     24.3 &     14.8 &       NA  \\
+median filter (2)    &    226.9 &       NA &   2034.0 &       NA  \\
+median filter (10)   &   2810.9 &       NA &   1877.1 &       NA  \\
+center mass          &      5.0 &       NA &   3611.2 &       NA  \\
+sobel                &     34.1 &       NA &     62.5 &      6.2  \\
+cwatershed           &    174.8 &  58440.3 &    287.3 &     44.9  \\
+daubechies           &     18.8 &       NA &       NA &       NA  \\
+haralick             &    233.1 &       NA &   7760.7 &       NA  \\
 \bottomrule
@@
 does not support. In the case of \emph{median filter}, the radius of the
-structuring element is shown in parentheses.}
-\label{tab:efficiency}
-\end{table}
+structuring element is shown in parentheses. NA stands for ``Not Available.''}
+\label{tab:efficiency} \end{table}
 
@@
 Linux. However, due to the normalization, measurements obtained on another
-system (Intel 32~bits running Mac OS) were qualitatively similar (data not
-shown).
+system (Intel 32~bits running Mac OS) were qualitatively similar.
 
@@
 interactive use as the user is often exploring and is willing to pay the speed
-cost of a few extra type conversions to avoid a hard-crash.
+cost of a few extra type checks and conversions to avoid a hard-crash.
 
 Among the other libraries, mahotas is the fastest. Pymorph, even though it is
-implemented in Python only, intelligently uses arithmetic operations for this
-operation and can be very fast. However, for more complex methods, such as
-watershed; its pure Python approach is very inefficient. The one exception is
-that median filtering with a large structuring element is faster in
-scikit-image. In fact, that library uses an algorithm with better asymptotic
-behavior for this operation.
+implemented in Python only, intelligently uses arithmetic operations for
+morphological operation and can be very fast. However, for more complex
+methods, such as watershed; its pure Python approach is very inefficient. The
+one exception is that median filtering with a large structuring element is
+faster in scikit-image. In fact, that library uses an algorithm with better
+asymptotic behavior for this operation.
 
-\bigskip
+\subsection{Distribution and Installation}
 
@@
 exist~\citep{Pedregosa:2011:SML:2078183.2078195,springerlink:10.1007/978-3-540-30116-5_58,Schaul:2010:PYB:1756006.1756030,Sonnenburg:2010:SML:1756006.1859911}.
-A good classification (or other) system can benefit both computer vision users
+A good classification system can benefit both computer vision users
 and others. As these projects all use Numpy arrays as their data types, it is
@@
 
-Python is an ideal language for fast development of both applications and
-scientific software. For this platform, Mahotas is a fast library of computer
-vision and image processing functions. It is implemented in \cpp{}, as the
-standard Python interpreter is too slow for a direct Python implementation.
-However, all of the Python interface code is hand-written, as opposed to using
-automatic interface generators like Swig~\citep{Beazley2003599}. This is more
-work intially, but the end result is of much higher quality, especially when it
-comes to giving useful error messages (e.g., when a type mismatch occurs, an
-automatic system will often be forced to resort to a generic message as it does
-not have any knowledge of what the arguments mean besides their automatically
-inferred types).
+Mahotas is implemented in \cpp{}, as the standard Python interpreter is too
+slow for a direct Python implementation. However, all of the Python interface
+code is hand-written, as opposed to using automatic interface generators like
+Swig~\citep{Beazley2003599}. This is more work initially, but the end result is
+of much higher quality, especially when it comes to giving useful error
+messages (e.g., when a type mismatch occurs, an automatic system will often be
+forced to resort to a generic message as it does not have any knowledge of what
+the arguments mean besides their automatically inferred types).
 
@@
 local binary patterns implementation. Finally, I thank the several users who
-have reported bugs, submitted small fixes, and participated on the project
+have reported bugs, submitted fixes, and participated on the project
 mailing list.
 
-Stéfan van der Walt and Andreas Müller offered helpful comments on a draft
+St\'efan van der Walt and Andreas M\"uller offered helpful comments on a draft
 version of this manuscript.

### 2013-03-20 paper.tex
REF Reference imglib2

--- 2013-01-02 paper.tex
@@
 computations. In particular, unlike libraries written in the C~Language or in
-Java~\citep{Marcel:2010:TMP:1873951.1874254}, Mahotas does not need to define a
+Java~\citep{Pietzsch15112012,Marcel:2010:TMP:1873951.1874254}, Mahotas does not need to define a
 new image data structure, but uses the numpy array structure. Many basic

### 2013-04-23 paper.tex
REF Add some missing references

@@
 \item[Features] Global feature descriptors. In particular, Haralick texture
-features, Zernike moments, local binary patterns, and threshold adjacency
-statistics (both the original~\citep{Hamilton2007} and the parameter-free
+features~\citep{Haralick1973}, Zernike moments, local binary
+patterns~\citep{Ojala2002}, and threshold adjacency statistics (both the
+original~\citep{Hamilton2007} and the parameter-free
 versions~\citep{Coelho2010}).
-\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms are
-supported.
+\item[Wavelet] Haar and Daubechies wavelets~\citep{Daubechies90}. Forward and
+inverse transforms are supported.
 \item[Morphological functions] Erosion and dilation, as well as some more

### 2013-04-23 paper.tex
MIN C++ compiler is needed. Update downloads
The download number is not yet at 50k, but close enough that it will
surely pass that mark before the paper is final.

@@
 
+For compiling from source, a \cpp{} compiler is needed, as well as the
+development headers for Python and numpy.
+
 \subsection{Quality Control}
@@
 Mahotas has been available in the Python Package Index since April~2010 and has
-been downloaded over 40,000~times. This does not include any downloads from
-other sources. Mahotas includes a full test suite. There are no known bugs.
+been downloaded over fifty thousand~times. This does not include any downloads
+from other sources. Mahotas includes a full test suite. There are no known
+bugs.
 

### 2013-04-23 paper.tex
REF New reference to work using mahotas

@@
 \textbf{Dependencies}\\
-It requires numpy to be present and installed.
+At runtime, mahotas requires numpy to be present and installed. If compiling
+from source, the Python headers and a \cpp{} compiler is required.
 
@@
 Originally, this code was developed in the context of cellular image analysis.
-However, none of the functionality is specific to this context and many
-computer vision pipelines can make use of it.
+However, the code was designed so that mahotas would contain functionality that
+is not specific to cell image analysis and many computer vision pipelines can
+make use of it.
 
-This package (and earlier versions of it) have been used by
+This package (or earlier versions of it) have been used by
 myself~\citep{Coelho2009,Coelho2010a} and close collaborators in several
-publications~\citep{omerosearcher}. Other groups have used in published work,
-both in cell image analysis~\citep{CYTO:CYTO22034} and in other
-areas~\citep{springerlink:10.1007/978-3-642-32335-5_2}.
+publications~\citep{omerosearcher}. Other groups have used it in published
+work, both in cell image analysis~\citep{CYTO:CYTO22034} and in other
+areas~\citep{springerlink:10.1007/978-3-642-32335-5_2,ADFM:ADFM201300091}.
 

### 2013-04-24 paper.tex
TXT Fix repetitions. Link to docs

@@
 available from \url{http://github.com/luispedro/mahotas} and from the Python Package
-Index (\url{http://pypi.python.org/pypi/mahotas}).
+Index (\url{http://pypi.python.org/pypi/mahotas}). Tutorials and full API
+documentation is available online at \url{http://mahotas.readthedocs.org/}.
 
@@
 well-tested package (the first versions were made available over 4~years ago,
-although the package was not named mahotas then). \footnote{Note for reviewers:
-the version currently available (0.9.6) implements all functionality described
-in this manuscript. I will release it as version 1.0 when this manuscript is
-accepted to coincide with publication. Naturally, any bugs that are found and
-reported in the meanwhile will be addressed.} Mahotas runs and is used on
+although the package was not named mahotas then). Mahotas runs and is used on
 different versions of Unix (including Linux, SunOS, and FreeBSD), Mac OS X, and
@@
 
+There are binary packages available for Windows, maintained by Christoph
+Gohlke, and for FreeBSD and Linux Frugalware through their respective package
+systems.
+
 \subsection{Quality Control}
@@
 \textbf{Additional system requirements}\\
-None at runtime. Compilation from source requires a \cpp{} compiler.
+None at runtime. Compilation from source requires a \cpp{} compiler and the
+Python development headers.
 
 \textbf{Dependencies}\\
-At runtime, mahotas requires numpy to be present and installed. If compiling
-from source, the Python headers and a \cpp{} compiler is required.
+At runtime, mahotas requires numpy to be present and installed.
 

### 2013-04-24 paper.tex
BUG Fix code bug in paper
Pointed out by reviewers

@@
 \begin{python}
-colors = np.array(
+colors = np.array([
     [ 255,  25,   1],

### 2013-04-26 paper.tex
ENH Replace milk by scipy.cluster.vq
This is a more natural dependency

--- 2013-04-24 paper.tex
@@
 We start by importing the necessary packages, including numpy and mahotas. We
-also use \textit{milk}, to demonstrate how the mahotas output can integrate
+also use \textit{scipy.cluster}, to demonstrate how the mahotas output can integrate
 with a machine learning package.
@@
 from mahotas.features import surf
-import milk
+from scipy.cluster import vq
 \end{python}
@@
 
-Using milk, we cluster the descriptors into five groups:
-
+Using \code{scipy.cluster.vq}, we cluster the descriptors into five groups. The
+function \code{kmeans2} returns two values: the centroids, which we ignore; and
+the cluster ids, which we will use below to assign colours:
 \begin{python}
-values, _ = milk.kmeans(descrs, 5)
+_,cids = vq.kmeans2(vq.whiten(descrs), 5)
 \end{python}
@@
     [ 47, 233,   5]])
-f2 = surf.show_surf(f, spoints[:64], values, colors)
+f2 = surf.show_surf(f, spoints[:64], cids, colors)
 \end{python}

### 2013-05-02 paper.tex
ENH Update date. Add another ref for mahotas use

@@
 Lane Center for Computational Biology, Carnegie Mellon University\\
-Instituto de Medicina Molecular}
-\date{January 2013}
+Instituto de Medicina Molecular\footnote{Current Address: European Molecular
+Biology Laboratory (EMBL), Heidelberg, Germany.}}
+
+\date{May 2013}
 
@@
 publications~\citep{omerosearcher}. Other groups have used it in published
-work, both in cell image analysis~\citep{CYTO:CYTO22034} and in other
-areas~\citep{springerlink:10.1007/978-3-642-32335-5_2,ADFM:ADFM201300091}.
+work in cell image analysis~\citep{CYTO:CYTO22034} and in other areas
+\citep{springerlink:10.1007/978-3-642-32335-5_2}. \citet{ADFM:ADFM201300091}
+used mahotas to detect nanoparticles in electron microscopy images.
+
+Mahotas provides many basic tools which can be combined to process images. It
+can be used in any problem which requires the processing of images to extract
+quantitative information.
 

### 2013-05-02 paper.tex
ENH Use mahotas.demos.image_path
This achieves file location independence

@@
 import numpy as np
-import mahotas
+import mahotas as mh
 from mahotas.features import surf
@@
 case, the conversion is done using standard numpy methods, namely
-\code{astype}:
+\code{astype}. We use the function \code{mahotas.demos.image\_path} to access a
+demonstration image.
 
 \begin{python}
-f = mahotas.imread('luispedro.jpg', as_grey=True)
+import mahotas.demos
+impath = mh.demos.image_path('luispedro.jpg')
+f = mh.imread(impath, as_grey=True)
 f = f.astype(np.uint8)

### 2013-05-03 paper.tex
ENH call pyplot.show() at end of demo
ENH Read through. Several corrections

@@
 and floating point images are supported by mahotas) can all be performed with
-numpy builtin functionality. For the user, this has the additional advantage
-that they do not need to learn yet another set of functions.
+numpy builtin functionality. For the users, this has the additional advantage
+that they do not need to learn new interfaces.
 
-It contains over 100~functions with functionality ranging from traditional
+Mahotas contains over 100~functions with functionality ranging from traditional
 image filtering and morphological operations to more modern wavelet
@@
 integer} or floating point numbers. While natural colour images are typically
-8~bits, scientific data is often larger and processing can result in floating
-point images. Mahotas works on all datatypes. This is performed without any
-extra memory copies. Mahotas is heavily optimised for both speed and memory
-usage (it can be used with very large arrays).
+8~bits, scientific data is often larger (12 and 16~bit formats are common).
+Processing can generate floating point images, for example, subtracting the
+mean value to every pixel; the mean will be a floating point number even if the
+original image was integral. Mahotas works on all datatypes. This is performed
+without any extra memory copies. Mahotas is heavily optimised for both speed
+and memory usage (it can be used with very large arrays).
 
@@
 We start by importing the necessary packages, including numpy and mahotas. We
-also use \textit{scipy.cluster}, to demonstrate how the mahotas output can integrate
-with a machine learning package.
+also use \code{scipy.cluster}, to demonstrate how the mahotas output can
+integrate with a machine learning package.
 
@@
 
-Finally, we can show the points in different colours.
+Finally, we can show the points in different colours. In order to avoid a very
+cluttered image, we will only plot the first 64~regions.
+
 \begin{python}
@@
 plt.imshow(f2)
+plt.show()
 \end{python}
@@
 The easy interaction with matplotlib is another way in which we benefit from
-the numpy-based ecosystem as mahotas does not need to support interacting with
-a graphical system to display images.
+the numpy-based ecosystem. Mahotas does not need to support interacting with a
+graphical system to display images.
 
@@
 generic one presented here. The selection between the two implementations is
-done automatically for the user.} \code{py\_erode} consists mostly of
+done automatically.} \code{py\_erode} consists mostly of
 boiler-plate code:
@@
 implementation of Python, there is a global lock for many Python related
-functionality, which limits parallelism.} in its constructor and gets it back
+functionality, which limits parallelism. By releasing the \textsc{gil}, mahotas
+allows another thread to run concurrently.} in its constructor and gets it back
 in its destructor. Normally, the template function will release the
@@
 normalized to multiples of the time it takes to go over the image and find its
-maximum pixel value (this was done using \code{numpy.max(image)}). The
+maximum pixel value (using the expression \code{numpy.max(image)}). The
 measurements shown were obtained on an Intel 64~bit system, running Ubuntu
-Linux. However, due to the normalization, measurements obtained on another
-system (Intel 32~bits running Mac OS) were qualitatively similar.
+Linux. Due to the normalization, measurements obtained on another system (Intel
+32~bits running Mac OS) were qualitatively similar.
 
@@
 interactive use as the user is often exploring and is willing to pay the speed
-cost of a few extra type checks and conversions to avoid a hard-crash.
+cost of a few extra type checks to avoid a hard-crash.
 
@@
 functionality and include several regression tests. There are no known bugs in
-version~1.0. In fact, no releases have ever been performed with known bugs.
+version~1.0. In fact, no releases have every contained any known bugs.
 Naturally, bugs were, occasionally, discovered in released versions, but
@@
 of much higher quality, especially when it comes to giving useful error
-messages (e.g., when a type mismatch occurs, an automatic system will often be
-forced to resort to a generic message as it does not have any knowledge of what
-the arguments mean besides their automatically inferred types).
+messages. When a type mismatch occurs, an automatic system will often be forced
+to resort to a generic message as it does not have any knowledge of what the
+arguments mean to the user. It will only know their automatically inferred
+types. A hand written system can also automatically convert arguments when
+meaningful and be more flexible without completely losing type checking.
 
@@
 \textsc{surf} implementation is a port from the code from
-\textit{dlib},\footnote{Dlib's webpage is at \url{http://dlib.net}.} a very
-good \cpp{} library by Davis King. I also gleaned some insight into the
+\textit{dlib},\footnote{Dlib's webpage is at \url{http://dlib.net}.} an
+excellent \cpp{} library by Davis King. I also gleaned some insight into the
 implementation of these features from Christopher Evan's OpenSURF library and
@@
 \textbf{Funding}: I was supported in my work by the Funda\c c\~{a}o para a
-Ci\^encia e Tecnologia (grants SFRH/BD/37535/2007 and
-PTDC/SAU-GMG/115652/2008), by NIH grant GM078622 (to Robert F.\ Murphy), by a
-grant from the Scaife Foundation, by the HHMI Interfaces Initiative, and by a
-grant from the Siebel Scholars Foundation.
+Ci\^encia e Tecnologia (grant SFRH/BD/37535/2007 to himself and grant
+PTDC/SAU-GMG/115652/2009 to Musa Mhlanga), by NIH grant GM078622 (to Robert F.\
+Murphy), by a grant from the Scaife Foundation, by the HHMI Interfaces
+Initiative, and by a grant from the Siebel Scholars Foundation.
 

### 2013-05-06 paper.tex
MIN Read through

@@
 Index (\url{http://pypi.python.org/pypi/mahotas}). Tutorials and full API
-documentation is available online at \url{http://mahotas.readthedocs.org/}.
+documentation are available online at \url{http://mahotas.readthedocs.org/}.
 
@@
 8~bits, scientific data is often larger (12 and 16~bit formats are common).
-Processing can generate floating point images, for example, subtracting the
-mean value to every pixel; the mean will be a floating point number even if the
+Processing can generate floating point images. For example, a common
+normalization procedure is to subtract from each pixel location the overall
+pixel value mean; the result will be a floating point image even if the
 original image was integral. Mahotas works on all datatypes. This is performed
@@
 Mahotas is mostly written in \cpp, but this is completely hidden from the user
-as there are hand-written Python wrappers for all functions. Automatically
-generated wrappers inevitably lead to worse error messages and are less
-flexible.
-
-The main reason that mahotas is implemented in \cpp{} (and not in C, which is
-the language of the Python interpreter) is to use templates. Almost \cpp{}
-functionality is split across 2~functions:
+as there are hand-written Python wrappers for all functions. The main reason
+that mahotas is implemented in \cpp{} (and not in C, which is the language of
+the Python interpreter) is to use templates. Almost all \cpp{} functionality is
+split across 2~functions:
 
@@
 types. A hand written system can also automatically convert arguments when
-meaningful and be more flexible without completely losing type checking.
+meaningful and be more flexible without completely foregoing type checking.
 

### 2013-05-06 figures/surf-tutorial.py
ENH Update code to match paper exactly

@@ -1,26 +1,28 @@
 import numpy as np
-import mahotas
+import mahotas as mh
 from mahotas.features import surf
-from os import path
-from pylab import *
-import milk
-
-f = mahotas.imread('luispedro.jpg', as_grey=True)
+from scipy.cluster import vq
+import mahotas.demos
+impath = mh.demos.image_path('luispedro.jpg')
+f = mh.imread(impath, as_grey=True)
 f = f.astype(np.uint8)
 spoints = surf.surf(f, 4, 6, 2)
 descrs = spoints[:,6:]
-
-values, _ = milk.kmeans(descrs, 5)
-colors = np.array([(255-52*i,25+52*i,37**i % 101) for i in xrange(5)])
-
-f2 = surf.show_surf(f, spoints[:64], values, colors)
-
-subplot(1,2,1)
-imshow(np.dstack([f,f,f]))
-xticks([])
-yticks([])
-subplot(1,2,2)
-imshow(f2)
-xticks([])
-yticks([])
-savefig('surf-tutorial.png')
+_,cids = vq.kmeans2(vq.whiten(descrs), 5)
+colors = np.array([
+    [ 255,  25,   1],
+    [203,  77,  37],
+    [151, 129,  56],
+    [ 99, 181,  52],
+    [ 47, 233,   5]])
+f2 = surf.show_surf(f, spoints[:64], cids, colors)
+from matplotlib import pyplot as plt
+plt.subplot(1,2,1)
+plt.imshow(np.dstack([f,f,f]))
+plt.xticks([])
+plt.yticks([])
+plt.subplot(1,2,2)
+plt.imshow(f2)
+plt.xticks([])
+plt.yticks([])
+plt.savefig('surf-tutorial.png')
